---
title: "An anti-inflammatory response of an organic food intervention treatment by reducing pesticide exposures in children of Cyprus: a cluster-randomized crossover trial - Sensitivity analysis Part 2"
output:
  html_document:
    toc: yes
    toc_float: yes
  word_document:
    toc: yes
authors: Cyprus International Institute for Environmental and Public Health, Cyprus
  University of Technology, Limassol, Cyprus.
---

# Sensitivity analysis Part 2

This script includes the sensitivity analysis (Part 2), in which samples from participants who reported medication use and/or illness within two days before sampling were excluded.
The output of the Part 2 script needs to be created (produced_data/lmmdata_dates_MOD).

```{r libraries, include=FALSE}

## Prepare workspace & install libraries

rm(list = ls(all = TRUE))

ipak <- function(pkg) {
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg)) {
    install.packages(new.pkg, dependencies = TRUE)
  }
  sapply(pkg, require, character.only = TRUE)
}

# usage
packages <- c(
  "readxl", "plyr", "dtplyr", "data.table", "tidyr", "knitr",
  "tibble", "Hmisc", "tableone", "compare", "stargazer",
  "readr", "ggplot2", "sjPlot", "tidyverse", "lubridate",
  "eeptools", "Rcpp", "nlme", "lme4", "lmerTest",
  "texreg", "reshape", "scales", "FSA", "NADA",
  "childsds", "sjPlot","ggpubr"
)

ipak(packages)

```

```{r linear mixed effect models - prepare database, echo=FALSE, include=FALSE}

# read the output of part 2
dir_path <- getwd()
lmmdata_dates_MOD <- readRDS(paste0(dir_path, "/produced_data/lmmdata_dates_MOD.rds", sep = ""))

# read file with samples that need to be excluded
excluded<-read.csv(file="Data//excluded.csv", header=T)

# create sample_code variable so that both dataframes have same variables
lmmdata_dates_MOD$samplecode <- paste(lmmdata_dates_MOD$Code, lmmdata_dates_MOD$sample_no_orig)
lmmdata_dates_MOD$sample_code <- gsub(' ','',lmmdata_dates_MOD$samplecode)

# create the dataset for the sensitivity analysis
lmmdata_dates_MOD_sens=lmmdata_dates_MOD[-which(lmmdata_dates_MOD$sample_code %in% excluded$Samples_List),]

# 24 samples are excluded

# vector with the titles of the rows used in the model specifications
rowtitles <- c(
  "Number of measurements",
  "Number of participants",
  "Number of schools"
)

```

## Models details

All models were adjusted for the baseline value of the outcome.

All models were adjusted for the creatinine levels using two ways:

*  CRP and pesticides measurements were divided with the respective creatinine measurements
*  Creatinine was included as a fixed effect

Initial model (model for objective 1):

 + Random part: participants
 + Fixed effect: treatment condition (organic or conventional) and time (days of treatment, where time=0 was the start of the treatment)
 + Interaction term: time*phase
*  Check interaction term:
  + if p-value>0.05 remove from the model
  
### Log creatinine adjusted CRP ~ baseline log creatinine adjusted CRP + DaysfromTreatment + phase + (phase*DaysfromTreatment if p<0.05) + random(...)

__Initial model (including the interaction term)__

```{r  crpmodel_INT, echo=FALSE, include=FALSE}

# model with interaction term
model_crp <- lmer(logadj_crp_S_TR ~ logadj_crp_S_BL + DaysfromTreatment + phase + DaysfromTreatment*phase + (1 | ID), lmmdata_dates_MOD_sens, na.action = na.exclude)

# summary(model_crp)
coefs_crp_INT <- merge(
  as.data.frame(
    round(
      # getting from the model coefficients table
      # the fixed effects names --- row.names
      # the estimate --- col. 1
      # the std. error --- col. 2
      # the p-value --- col. 5
      coef(summary(model_crp))[, c(1, 2, 5)],
      digits = 3
    )
  ),
  as.data.frame(
    round(
      # getting the confidence intervals omitting the first 3 rows that
      # refer to sigma (it is NA in the Wald method)
      confint.merMod(model_crp, method = "Wald"),
      digits = 3
    )[-c(1:3), ]
  ),
  by = "row.names"
)

# change the names in the columns of the CIs
setnames(coefs_crp_INT, old = c("Pr(>|t|)", "2.5 %", "97.5 %"), new = c("p", "CI2.5", "CI97.5"))

coefs_crp_INT$Model <- "crpadjusted_INT"

# revalue the fixed effects names
coefs_crp_INT$FixedEffects <- revalue(
  coefs_crp_INT$Row.names,
  c(
    #"(Intercept)" = "Intercept",
    "DaysfromTreatment" = "Time (days)",
    "DaysfromTreatment:phaseOrganic"="Days*Phase",
    "phaseOrganic" = "Organic diet",
    "logadj_crp_S_BL" = "Baseline CRP levels"
  )
)

model_specs_crp <- c(
  # Number of measurements taken from the length of the outcome variable used in the model
  length(model_crp@frame$logadj_crp_S_TR),
  # Number of participants taken from the levels of the ID variables
  nlevels(model_crp@frame$ID),
  # Number of schools taken from the model details
  nlevels(model_crp@flist$SCID)
)

# add the row titles to the model specs
obs_dets_crp <- data.frame(rowtitles, model_specs_crp)
names(obs_dets_crp) <- c("Model details", "")

```

```{r tabmpdel Model 1, echo=FALSE, include=TRUE}
# crp
tab_model(model_crp,
  show.icc = TRUE,
  digits = 3, show.obs = TRUE, show.re.var = TRUE,
  title = paste(model_crp@call$formula[3]))
# model has been inspected: interaction term p-value<0.05

```

```{r model_crp, echo=FALSE}

plot(model_crp, resid(.) ~ fitted(.) | phase)

plot(model_crp, resid(.) ~ DaysfromTreatment)

qqnorm(resid(model_crp))

```

### Log CRP ~ baseline Log CRP + DaysfromTreatment + phase + creatinine_S + (phase*DaysfromTreatment if p<0.05) + random(...)

```{r  crpmodel_INT_logcr_int, echo=FALSE, include=FALSE}

# model with interaction term
model_crp_cr_INT <- lmer(logcrp_S_TR ~ logcrp_S_BL + DaysfromTreatment + phase + creatinine_S + phase*DaysfromTreatment +(1 | ID), lmmdata_dates_MOD_sens, na.action = na.exclude)

```

```{r tabmodel Model logcr int, echo=FALSE, include=TRUE}
# crp

#tab_model(model_crp_cr_INT,
#  show.icc = TRUE,
#  digits = 3, show.obs = TRUE, show.re.var = TRUE,
#  title = paste(model_crp_cr_INT@call$formula[3]))
# model has been inspected: interaction term p-value>0.05

```

```{r model_crp_logcr_INT, echo=FALSE}

#plot(model_crp_cr_INT, resid(.) ~ fitted(.) | phase)

#plot(model_crp_cr_INT, resid(.) ~ DaysfromTreatment)

#qqnorm(resid(model_crp_cr_INT))

```


```{r  crpmodel_cr_nonINT , echo=FALSE, include=FALSE}

# model without interaction term
model_crp_cr_nonINT <- lmer(logcrp_S_TR ~ logcrp_S_BL + DaysfromTreatment + phase + creatinine_S +(1 | ID), lmmdata_dates_MOD_sens, na.action = na.exclude)

# summary(model_crp_cr_nonINT)
coefs_crp_cr_nonINT <- merge(
  as.data.frame(
    round(
      # getting from the model coefficients table
      # the fixed effects names --- row.names
      # the estimate --- col. 1
      # the std. error --- col. 2
      # the p-value --- col. 5
      coef(summary(model_crp_cr_nonINT))[, c(1, 2, 5)],
      digits = 3
    )
  ),
  as.data.frame(
    round(
      # getting the confidence intervals omitting the first 3 rows that
      # refer to sigma (it is NA in the Wald method)
      confint.merMod(model_crp_cr_nonINT, method = "Wald"),
      digits = 3
    )[-c(1:3), ]
  ),
  by = "row.names"
)

# change the names in the columns of the CIs
setnames(coefs_crp_cr_nonINT, old = c("Pr(>|t|)", "2.5 %", "97.5 %"), new = c("p", "CI2.5", "CI97.5"))

coefs_crp_cr_nonINT$Model <- "crp_nonadjusted_nonINT"

# revalue the fixed effects names
coefs_crp_cr_nonINT$FixedEffects <- revalue(
  coefs_crp_cr_nonINT$Row.names,
  c(
    "(Intercept)" = "Intercept",
    "DaysfromTreatment" = "Time (days)",
    "phaseOrganic" = "Organic diet",
    "logcrp_S_BL" = "Baseline crp levels (no cr adjusted)",
    "creatinine_S"="creatinine"  
  )
)

```

```{r tabmodel Model cr, echo=FALSE, include=TRUE}
# crp
tab_model(model_crp_cr_nonINT,
  show.icc = TRUE,
  digits = 3, show.obs = TRUE, show.re.var = TRUE,
  title = paste(model_crp_cr_nonINT@call$formula[3]))

```

```{r model_crp_cr_nonINT, echo=FALSE}

plot(model_crp_cr_nonINT, resid(.) ~ fitted(.) | phase)

plot(model_crp_cr_nonINT, resid(.) ~ DaysfromTreatment)

qqnorm(resid(model_crp_cr_nonINT))

```

### Merge the models of objective 1

```{r tabmodel merge the models of objective 1, echo=FALSE, include=TRUE}
# crp
tab_model(model_crp,model_crp_cr_nonINT,
  show.icc = TRUE,
  digits = 3, show.obs = TRUE, show.re.var = TRUE,
  title = "Merge the models of objective 1")

```

### Models for objective 2

A second model was made to assess the association between biomarker of exposure to pesticides (3-PBA and 6-CN, separately) and biomarker of inflammation (CRP). 

 + Random part: participants
 + Fixed effect: each biomarker of exposure to pesticides, time (days of treatment, where time=0 was the start of the treatment), age and sex


### Log creatinine adjusted CRP ~ Log creatinine adjusted baseline CRP + 6-CN (binary) + DaysfromTreatment + sex + baseline age + random(...)

```{r add 6-CN, echo=FALSE, include=TRUE}
# crp

Model2<-lmer(logadj_crp_S_TR ~ logadj_crp_S_BL + cnlod + DaysfromTreatment + Sex + age_baseline_S + (1 | ID), lmmdata_dates_MOD_sens, na.action = na.exclude)

# summary(model_crp_noINT)

coefs_model2_noINT <- merge(
  as.data.frame(
    round(
      # getting from the model coefficients table
      # the fixed effects names --- row.names
      # the estimate --- col. 1
      # the std. error --- col. 2
      # the p-value --- col. 5
      coef(summary(Model2))[, c(1, 2, 5)],
      digits = 3
    )
  ),
  as.data.frame(
    round(
      # getting the confidence intervals omitting the first 3 rows that
      # refer to sigma (it is NA in the Wald method)
      confint.merMod(Model2, method = "Wald"),
      digits = 3
    )[-c(1:3), ]
  ),
  by = "row.names"
)

# change the names in the columns of the CIs
setnames(coefs_model2_noINT, old = c("Pr(>|t|)", "2.5 %", "97.5 %"), new = c("p", "CI2.5", "CI97.5"))

coefs_model2_noINT$Model <- "crpadjusted_6CN"

# revalue the fixed effects names
coefs_model2_noINT$FixedEffects <- revalue(
  coefs_model2_noINT$Row.names,
  c(
    "cnlodAbove LOD"="6-CN (above LOD)",
    "logadj_crp_S_BL"="Baseline CRP levels",
    "DaysfromTreatment" = "Time (days)",
     "SexFemale"=="Sex (female)",
    "age_baseline_S"="Age"
  )
)
```

```{r echo=FALSE, include=TRUE}
tab_model(Model2,
  show.icc = TRUE,
  digits = 3, show.obs = TRUE, show.re.var = TRUE,
  title = paste(Model2@call$formula[3])
)
```

```{r diagnostics_model2, echo=FALSE}

plot(Model2, resid(.) ~ fitted(.) | phase)

plot(Model2, resid(.) ~ DaysfromTreatment)

qqnorm(resid(Model2))

```

### Log CRP ~ log baseline CRP + 6-CN + DaysfromTreatment + sex + baseline age + creatinine_s + random(...)

```{r add 6-CN non adjusted, echo=FALSE, include=TRUE}
# crp

Model2b<-lmer(logcrp_S_TR ~ logcrp_S_BL + cnlod + DaysfromTreatment + Sex + age_baseline_S + creatinine_S + (1 | ID), lmmdata_dates_MOD_sens, na.action = na.exclude)

# summary(model_crp_noINT)

coefs_Model2b_noINT <- merge(
  as.data.frame(
    round(
      # getting from the model coefficients table
      # the fixed effects names --- row.names
      # the estimate --- col. 1
      # the std. error --- col. 2
      # the p-value --- col. 5
      coef(summary(Model2b))[, c(1, 2, 5)],
      digits = 3
    )
  ),
  as.data.frame(
    round(
      # getting the confidence intervals omitting the first 3 rows that
      # refer to sigma (it is NA in the Wald method)
      confint.merMod(Model2b, method = "Wald"),
      digits = 3
    )[-c(1:3), ]
  ),
  by = "row.names"
)

# change the names in the columns of the CIs
setnames(coefs_Model2b_noINT, old = c("Pr(>|t|)", "2.5 %", "97.5 %"), new = c("p", "CI2.5", "CI97.5"))

coefs_Model2b_noINT$Model <- "crp_nonadjusted_6CN"

# revalue the fixed effects names
coefs_Model2b_noINT$FixedEffects <- revalue(
  coefs_Model2b_noINT$Row.names,
  c(
    "cnlodAbove LOD"="6CN (above LOD)",
    "logcrp_S_BL"="Baseline crp levels (no cr adjusted)",
    "DaysfromTreatment" = "Time (days)",
     "SexFemale"=="Sex (female)",
    "age_baseline_S"="Age",
    "creatinine_S"="creatinine" 
    
  )
)

```

```{r echo=FALSE, include=TRUE}
tab_model(Model2b,
  show.icc = TRUE,
  digits = 3, show.obs = TRUE, show.re.var = TRUE,
  title = paste(Model2b@call$formula[3])
)

```

```{r diagnostics_Model2b, echo=FALSE}

plot(Model2b, resid(.) ~ fitted(.) | phase)

plot(Model2b, resid(.) ~ DaysfromTreatment)

qqnorm(resid(Model2b))

```

### Merge the models of objective 2-6-CN

```{r merge model of objective 2 6-cn, echo=FALSE, include=TRUE}

tab_model(Model2,Model2b,
  show.icc = TRUE,
  digits = 3, show.obs = TRUE, show.re.var = TRUE,
  title = "Merge the models of objective 2-6-CN")

```

### Log creatinine adjusted CRP ~ Log creatinine adjusted baseline CRP + creatinine adjusted 3-PBA + DaysfromTreatment + sex + baseline age + random(...)

```{r add 3-PBA, echo=FALSE, include=TRUE}
# PBA

Model3<-lmer(logadj_crp_S_TR ~ logadj_crp_S_BL + adj_pb1000_S + DaysfromTreatment + Sex + age_baseline_S + (1 | ID), lmmdata_dates_MOD_sens, na.action = na.exclude)

# summary(model_crp_noINT)

coefs_model3_noINT <- merge(
  as.data.frame(
    round(
      # getting from the model coefficients table
      # the fixed effects names --- row.names
      # the estimate --- col. 1
      # the std. error --- col. 2
      # the p-value --- col. 5
      coef(summary(Model3))[, c(1, 2, 5)],
      digits = 3
    )
  ),
  as.data.frame(
    round(
      # getting the confidence intervals omitting the first 3 rows that
      # refer to sigma (it is NA in the Wald method)
      confint.merMod(Model3, method = "Wald"),
      digits = 3
    )[-c(1:3), ]
  ),
  by = "row.names"
)

# change the names in the columns of the CIs
setnames(coefs_model3_noINT, old = c("Pr(>|t|)", "2.5 %", "97.5 %"), new = c("p", "CI2.5", "CI97.5"))

coefs_model3_noINT$Model <- "crpadjusted_3PBA"

# revalue the fixed effects names
# revalue the fixed effects names
coefs_model3_noINT$FixedEffects <- revalue(
  coefs_model3_noINT$Row.names,
  c(
    "adj_pb1000_S"="3PBA",
    "logadj_crp_S_BL"="Baseline crp levels",
    "DaysfromTreatment" = "Time (days)",
     "SexFemale"=="Sex (female)",
    "age_baseline_S"="Age"
  )
)

```
```{r echo=FALSE, include=TRUE}
tab_model(Model3,
  show.icc = TRUE,
  digits = 3, show.obs = TRUE, show.re.var = TRUE,
  title = paste(Model3@call$formula[3])
)

```

```{r diagnostics_model3, echo=FALSE}

plot(Model3, resid(.) ~ fitted(.) | phase)

plot(Model3, resid(.) ~ DaysfromTreatment)

qqnorm(resid(Model3))

```

### Log CRP ~ Log baseline CRP + 3-PBA + DaysfromTreatment + sex + baseline age + creatinine_S + random(...)

```{r add 3-PBA non adjusted, echo=FALSE, include=TRUE}
# PBA

Model3b<-lmer(logcrp_S_TR ~ logcrp_S_BL + pb1000_S + DaysfromTreatment + Sex + age_baseline_S + creatinine_S + (1 | ID), lmmdata_dates_MOD_sens, na.action = na.exclude)

# summary(model_crp_noINT)

coefs_Model3b_noINT <- merge(
  as.data.frame(
    round(
      # getting from the model coefficients table
      # the fixed effects names --- row.names
      # the estimate --- col. 1
      # the std. error --- col. 2
      # the p-value --- col. 5
      coef(summary(Model3b))[, c(1, 2, 5)],
      digits = 3
    )
  ),
  as.data.frame(
    round(
      # getting the confidence intervals omitting the first 3 rows that
      # refer to sigma (it is NA in the Wald method)
      confint.merMod(Model3b, method = "Wald"),
      digits = 3
    )[-c(1:3), ]
  ),
  by = "row.names"
)

# change the names in the columns of the CIs
setnames(coefs_Model3b_noINT, old = c("Pr(>|t|)", "2.5 %", "97.5 %"), new = c("p", "CI2.5", "CI97.5"))

coefs_Model3b_noINT$Model <- "crp_nonadjusted_6PBA"

# revalue the fixed effects names
# revalue the fixed effects names
coefs_Model3b_noINT$FixedEffects <- revalue(
  coefs_Model3b_noINT$Row.names,
  c(
    "pb1000_S"="3-PBA",
    "log_crp_S_BL"="Baseline CRP levels (no cr adjusted)",
    "DaysfromTreatment" = "Time (days)",
     "SexFemale"=="Sex (female)",
    "age_baseline_S"="Age",
    "creatinine_S"="creatinine"
  )
)

```

```{r echo=FALSE, include=TRUE}

tab_model(Model3b,
  show.icc = TRUE,
  digits = 3, show.obs = TRUE, show.re.var = TRUE,
  title = paste(Model3b@call$formula[3])
)
```

```{r diagnostics_Model3b, echo=FALSE}

plot(Model3b, resid(.) ~ fitted(.) | phase)

plot(Model3b, resid(.) ~ DaysfromTreatment)

qqnorm(resid(Model3b))

```

### Merge the models of objective 2-3-PBA

```{r merge the models of objective 2-3-pba, echo=FALSE, include=TRUE}

tab_model(Model3,Model3b,
  show.icc = TRUE,
  digits = 3, show.obs = TRUE, show.re.var = TRUE,
  title = "Merge the models of objective 2 3-BPA")

```

## Benjamini-Hochberg p-value adjustment - without the intercepts

```{r pvalueadj_all2, echo=FALSE, include=FALSE}

# combine the p-values of all the models made so far
pvalmod2 <- list(
  coefs_crp_INT[, c("Model", "FixedEffects", "p")],
  coefs_crp_cr_nonINT[, c("Model", "FixedEffects", "p")],
  coefs_model2_noINT[, c("Model", "FixedEffects", "p")],
  coefs_Model2b_noINT[, c("Model", "FixedEffects", "p")],
  coefs_model3_noINT[, c("Model", "FixedEffects", "p")],
  coefs_Model3b_noINT[, c("Model", "FixedEffects", "p")]
  ) # 2

all_pvalmod2 <- bind_rows(pvalmod2)

### Order data by p-value
all_pvalmod2 <- all_pvalmod2[order(all_pvalmod2$p), ]

# Benjamini-Hochberg
all_pvalmod2$BH_p <- round(p.adjust(all_pvalmod2$p, method = "BH", n = 30), digits = 3)

```

`r kable(all_pvalmod2)`

# Session information

```{r session, echo=FALSE}
sessionInfo()
```
