---
title: "An anti-inflammatory response of an organic food intervention treatment by reducing pesticide exposures in children of Cyprus: a cluster-randomized crossover trial - Part 1"
output:
  html_document: default
  word_document: default
authors: Cyprus International Institute for Environmental and Public Health, Cyprus
  University of Technology, Limassol, Cyprus.
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
  
The script was adapted by the script prepared by Corina Konstantinou and Xanthi Andrianou for the statistical analysis of the cluster-randomized crossover trial of organic diet for primary school children, conducted within the EU LIFE+ programme funded project Organiko in Cyprus (Makris et al., 2019).

This script includes Part 1 of the analysis (descriptives, description of the change in biomarker levels) and the dataset that will be the input of Part 2 of the analysis.

```{r libraries, include=FALSE}

## Prepare workspace & install libraries
rm(list = ls(all = TRUE))

ipak <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg)) 
    install.packages(new.pkg, dependencies = TRUE)
  sapply(pkg, require, character.only = TRUE)
}

# usage
packages <- c("readxl","plyr","dtplyr","data.table","tidyr","knitr",
              "tibble","Hmisc","tableone","compare","stargazer",
              "readr","ggplot2","sjPlot","tidyverse","lubridate",
              "eeptools", "Rcpp", "nlme", "lme4","lmerTest",
              "texreg","reshape","scales","FSA","NADA",
              "childsds", "sjPlot")

ipak(packages)

```

```{r baseline data reading and inspection, include=FALSE}
## Baseline questionnaire - Data inspection

#Three files named:
#- Baseline_Ques(Final) 250517-EA
#- Baseline_Ques(Final) 250517-SI
#- Baseline_Ques(Final) 250517-IC are located in the "Analysis" folder.

#The compiled data file "mergedbaseline" contains 191 observations and 
#72 variables. 

#Variable names, description, type and valid values are described in the file:
#"Variables characteristics" located in the same folder.

#The below actions have been done for data inspection:

#- Read data
#- Merge the 3 files into 1
#- Create school variable by deleting last 2 characters of Code
#- Change variable types into factors/characters
#- Format dates

#Additional actions:

#- Delete variable V20 because it contains no data 
#- Convert 555 (missing values) to NA
#- Calculate age at baseline

#+ first subset in order to omit NA
#+ merge baseline data with age data

#-Merge baseline data with samples per participant data. The file "Samples per participant.csv" 
#is located in the "Analysis" folder and contains 191 observations and 7 variables. Variable names, description, type and valid values are 
#described in the file "Variables characteristics" located in the same folder.

# read data & merge 
baseline1<-read.csv(file="Data//Baseline_Ques(Final) 250517-EA.csv", 
header=T, sep=";")
baseline2<-read.csv(file="Data//Baseline_Ques(Final) 250517-SI.csv", 
header=T, sep=";")
baseline3<-read.csv(file="Data//Baseline_Ques(Final) 250517-IC.csv", 
header=T, sep=";")

baseline<-rbind.data.frame(baseline1, baseline2, baseline3)

# merge baseline data with samples per participant data
samplesperpart<-read.csv(file="Data//Samples and days per participant.csv", header=T, sep=";")
# check variable types
str(samplesperpart)

# change code to factor
samplesperpart$Code<- as.factor(samplesperpart$Code)

mergedbaseline<-merge(baseline,samplesperpart, by="Code")

```

```{r recode variables, include=FALSE}
# create school variable by deleting last 2 characters of Code 
# Code needs to be a character
mergedbaseline$Code<-as.character(mergedbaseline$Code)
mergedbaseline$school<- substr(mergedbaseline$Code, 
                               start=1,
                               stop=nchar(mergedbaseline$Code)-2)


# change variable types into factors
mergedbaseline[, c("Editor", "Code", "Group", "Completion_study",
                   "medulev", "Other_educational_level_m",
                   "fedulev", "Gender", "handinmouth", "school")]<-lapply(
                     mergedbaseline[, c("Editor", "Code", "Group", 
                                        "Completion_study",
                                        "medulev", "Other_educational_level_m",
                                        "fedulev", "Gender", "handinmouth",
                                        "school")], factor)

# change variable types into characters
mergedbaseline[, c("city_village", "product_name",
                   "other_activity_name", "Other_Ctivity_name_1",
                   "other_sedent_name", "Other_sedent_name_1")]<- lapply(
                     mergedbaseline[, c("city_village",
                                        "product_name", "other_activity_name",
                                        "Other_Ctivity_name_1",
                                        "other_sedent_name", 
                                        "Other_sedent_name_1")], as.character)

# format dates 
mergedbaseline[, c("Qcompletion_date", "birthdate", "anthr_date_start",
                   "anthr_date_endorg","anthr_date_endconv")]<-lapply(
                     mergedbaseline[, c("Qcompletion_date", "birthdate",
                                        "anthr_date_start",
                                        "anthr_date_endorg",
                                        "anthr_date_endconv")],
                     as.Date, format="%d/%m/%Y")

# delete variable V20 because it contains no data
which(!is.na(mergedbaseline$V20))
#16  29 180 184 186 189 191)
mergedbaseline1<-subset(mergedbaseline, select=-V20) 

# convert 555 (missing values) to NA
mergedbaseline1[mergedbaseline1==555] <- NA
str(mergedbaseline1)

# create total samples number variable by adding all separate samples
mergedbaseline1$totsamples<-with(mergedbaseline1, S1 + S2 + S3 + S4 + S5 + S6)

# change totsamples variable  into factors
mergedbaseline1$totsamples<-as.factor(mergedbaseline1$totsamples)

# calculate age with decimals
# calculate the age at the start of the study
# subset with date start and birthdate in order to omit NA and calculate age_start
# Make NA the birthdate "1900-01-05" 
mergedbaseline1[mergedbaseline1=="1900-01-05"] <- NA
baseline_age1<-subset(mergedbaseline1, 
                      select=c("Code","birthdate","anthr_date_start"))
baseline_age1 <- na.omit(baseline_age1)
baseline_age1$age_baseline<-with(baseline_age1, 
                                 eeptools::age_calc(dob=birthdate, 
                                          enddate=anthr_date_start, 
                                          units = "years"))

mergedbaseline_age1<-merge(mergedbaseline1, baseline_age1[c("Code","age_baseline")], by="Code",
                           all = TRUE)

# calculate the age at the end of the organic treatment
baseline_age2<-subset(mergedbaseline1, 
                      select=c("Code","birthdate","anthr_date_endorg"))
baseline_age2 <- na.omit(baseline_age2)
baseline_age2$age_endorg<-with(baseline_age2, 
                               eeptools::age_calc(dob=birthdate,
                                        enddate=anthr_date_endorg, 
                                        units = "years"))

mergedbaseline_age2<-merge(mergedbaseline_age1, 
                           baseline_age2[c("Code","age_endorg")], by="Code",all = TRUE)

# calculate the age at the end of the conventional treatment for Group 1 only
baseline_age3<-subset(mergedbaseline1,
                      select=c("Code","birthdate","anthr_date_endconv"))
baseline_age3 <- na.omit(baseline_age3)
baseline_age3$age_endcon<-with(baseline_age3, 
                               age_calc(dob=birthdate,
                                        enddate=anthr_date_endconv, 
                                        units = "years"))

mergedbaseline2<-merge(mergedbaseline_age2,
                       baseline_age3[c("Code","age_endcon")], by="Code",all = TRUE)

# revalue categorical variables to their actual values
mergedbaseline2$Sex <- revalue(mergedbaseline2$Gender, 
                               c("1"="Male", "2"="Female"))

# calculate BMI based on formula: weight/height^2
mergedbaseline2$BMI_start<-mergedbaseline2$Weight_start/
                          ((mergedbaseline2$Height_start/100)^2)
mergedbaseline2$BMI_endorg<-mergedbaseline2$Weight_endorg/
                          ((mergedbaseline2$Height_endorg/100)^2)
mergedbaseline2$BMI_endcon<-mergedbaseline2$Weight_endconv/
                          ((mergedbaseline2$Height_endconv/100)^2)

# calculate BMI SDS based on WHO 2007 growth references
mergedbaseline2$BMI_start_sds <- sds(mergedbaseline2$BMI_start,
                       age = mergedbaseline2$age_baseline,
                       sex = mergedbaseline2$Sex, 
                       male = "Male", female = "Female",
                       ref = who2007.ref,
                       item = "bmi",
                       type = "SDS")

mergedbaseline2$BMI_endorg_sds <- sds(mergedbaseline2$BMI_endorg,
                       age = mergedbaseline2$age_endorg,
                       sex = mergedbaseline2$Sex, 
                       male = "Male", female = "Female",
                       ref = who2007.ref,
                       item = "bmi",
                       type = "SDS")

mergedbaseline2$BMI_endcon_sds <- sds(mergedbaseline2$BMI_endcon,
                       age = mergedbaseline2$age_endcon,
                       sex = mergedbaseline2$Sex, 
                       male = "Male", female = "Female",
                       ref = who2007.ref,
                       item = "bmi",
                       type = "SDS")

# Check the new BMI variables
summary(mergedbaseline2$BMI_start_sds)
summary(mergedbaseline2$BMI_endorg_sds)
summary(mergedbaseline2$BMI_endcon_sds)

# create categories based on WHO cut-off points for 
# thinness, normal, overweight and obese.
# <-2: Thinness, 1>-2: Normal, >1: Overweight, >2: Obese
mergedbaseline2$BMI_for_age_start <- cut(mergedbaseline2$BMI_start_sds,
                               breaks=c(-Inf, -2, 1, 2, Inf),
                               labels=c("Thinness","Normal","Overweight","Obese"))
mergedbaseline2$BMI_for_age_endorg <- cut(mergedbaseline2$BMI_endorg_sds,
                               breaks=c(-Inf, -2, 1, 2, Inf),
                               labels=c("Thinness","Normal","Overweight","Obese"))
mergedbaseline2$BMI_for_age_endcon <- cut(mergedbaseline2$BMI_endcon_sds,
                               breaks=c(-Inf, -2, 1, 2, Inf),
                               labels=c("Thinness","Normal","Overweight","Obese"))

# revalue categorical variables to their actual values: 
# For 6=Other, 7=Missing and 8=NA use NA
mergedbaseline2$Group1 <- revalue(mergedbaseline2$Group, 
                                  c("1"="First organic", "2"="First conventional"))
mergedbaseline2$medulev1 <- revalue(mergedbaseline2$medulev, 
                                    c("1"="Primary", "2"="Secondary", 
                                      "3"="University/college", 
                                      "4"="Master/PhD", "6"= NA,"7"=NA, "8"=NA))
mergedbaseline2$fedulev1 <- revalue(mergedbaseline2$fedulev, 
                                    c("1"="Primary", "2"="Secondary", 
                                      "3"="University/college", 
                                      "4"="Master/PhD", "6"=NA,"7"=NA, "8"=NA))
mergedbaseline2$Handinmouth1 <- revalue(mergedbaseline2$handinmouth, 
                                        c("0"="No", "1"="Yes","7"=NA))

# recode Days_Org in categories 0-11 days (dropouts), 12-40 days (active participants)
mergedbaseline2$daysorg <- cut(mergedbaseline2$Days_Org,
                               breaks=c(-Inf, 11, 40),
                               labels=c("Dropouts","Active participants"))

# recode Days_Org in categories 0-11, 12-21, 22-28, 29-40
mergedbaseline2$daysorgcat <- cut(mergedbaseline2$Days_Org,
                               breaks=c(-Inf, 11, 21, 28, 40),
                               labels=c("0-11 days","12-21 days","22-28 days","29-40 days"))

# Last make the variable selection
selectionDescr<-c("Group1","medulev1","fedulev1","Sex","age_baseline",
                  "Waist_start", "Waist_endorg","Waist_endconv",
                  "BMI_start","BMI_endorg","BMI_endcon",
                  "BMI_for_age_start","BMI_for_age_endorg","BMI_for_age_endcon",
                  "totsamples","daysorgcat","Handinmouth1")

categorvariablesDescr<-c("Group1","medulev1","fedulev1","Sex",
                         "BMI_for_age_start","BMI_for_age_endorg","BMI_for_age_endcon",
                         "totsamples","daysorgcat","Handinmouth1")

# convert NA values to 0 in columns where the times per week are denoted in order to be included in the stats
mergedbaseline2[c("park_week","garden_week",
                  "run_week", "cyc_week", "bask_week", "voll_week", "swim_week",
                  "danc_week", "other_activity_time","Other_activity_time_1",
                  "tv_week", "comp_week",
                  "tablet_week", "mobile_week",
                  "other_sedent_week","Other_sedent_week_1")][is.na(mergedbaseline2[c("park_week","garden_week",
                  "run_week", "cyc_week", "bask_week", "voll_week", "swim_week",
                  "danc_week", "other_activity_time","Other_activity_time_1",
                  "tv_week", "comp_week",
                  "tablet_week", "mobile_week",
                  "other_sedent_week","Other_sedent_week_1")])] <- 0

# summarize time spent outside
mergedbaseline2$outside_time<-with(mergedbaseline2, park_week+garden_week)

# summarize time spent for physical activities
mergedbaseline2$physactiv_time<-with(mergedbaseline2, run_week+cyc_week+
                                       bask_week+voll_week+swim_week+danc_week+
                                       other_activity_time+Other_activity_time_1)

# summarize time spent for sedentary activities
mergedbaseline2$sedentactiv_time<-with(mergedbaseline2, tv_week+comp_week+
                                                           tablet_week+mobile_week+
                                                           other_sedent_week+
                                                           Other_sedent_week_1)

selectionActivities<-c("garden_week","park_week","run_week","cyc_week",
                       "bask_week","foot_week","voll_week","swim_week",
                       "danc_week","tv_week","comp_week", "tablet_week",
                       "mobile_week","outside_time","physactiv_time",
                       "sedentactiv_time")

# create subset with participants with more than 12 days in the organic phase (active)
mergedbaseline3<-droplevels(mergedbaseline2[which(mergedbaseline2$Days_Org > 11), ])

# create subset with BMI data of the active participants per reviewer's request
bmidata<-mergedbaseline3[c("Code","BMI_start","BMI_endorg",
                  "BMI_start_sds", "BMI_endorg_sds", 
                  "BMI_for_age_start", "BMI_for_age_endorg")]

# write csv file with individual BMI data
BMI_ind_data<-write.csv(bmidata,file="produced_data/bmidata.csv", row.names=FALSE)

# check normality of variables
qqnorm(mergedbaseline3$Waist_start) 
hist(mergedbaseline3$Waist_start)

qqnorm(mergedbaseline3$physactiv_time)  
hist(mergedbaseline3$physactiv_time)

qqnorm(mergedbaseline3$sedentactiv_time) 
hist(mergedbaseline3$sedentactiv_time)

selectionnonnorm<-c("Waist_start", "Waist_endorg","Waist_endconv")
selectionactnonnorm<-c("physactiv_time","sedentactiv_time")

# create subset with participants with less than 12 days in the organic phase
mergedbaselinedropouts<-droplevels(mergedbaseline2[which(mergedbaseline2$Days_Org < 12), ])

# round age variables
mergedbaseline3$age_baseline<-round(mergedbaseline3$age_baseline)
mergedbaseline3$age_endorg<-round(mergedbaseline3$age_endorg)
mergedbaseline3$age_endcon<-round(mergedbaseline3$age_endcon)

```

### Table 1 A - Demographic characteristics - Study participants

`r kable(print(CreateTableOne(selectionDescr, data=mergedbaseline3, factorVars=categorvariablesDescr, testNonNormal = kruskal.test, smd=F), nonnormal=selectionnonnorm), caption="Demographic & other characteristics of participants with >=12 days in the organic phase")`

`r kable(print(CreateTableOne(selectionDescr, data=mergedbaseline3, factorVars=categorvariablesDescr, strata="Group1", testNonNormal = kruskal.test, smd=F), nonnormal=selectionnonnorm), caption="Demographic & other characteristics of participants with >=12 days in the organic phase by group")`

## Urinary data

```{r Pesticides and biomarkers, include=FALSE}

rm(list=setdiff(ls(),c("mergedbaseline3", "mergedbaseline2")))

# Read measurements

meas_mdacr<-read.csv(file="Data//mda_cr_meas.csv", header=T, sep=";")
str(meas_mdacr)
dim(meas_mdacr)
#941 3

#Keep the samples that the creatinine measurements are between 0.3 and 3 g/L
meas_mdacr_sp<-meas_mdacr%>%
 filter(creatinine>0.29)%>%
  filter(creatinine<3.01)
dim(meas_mdacr_sp)
#927 3

#There are 941-927=14 samples withcreatinine measurements lower than 0.3 and higher than 3 g/L

meas_pest<-read.csv(file="Data//Pesticides_meas.csv", header=T, sep=";")
str(meas_pest)
dim(meas_pest)
#941 3

#Units CRP:ng/L
meas_crpr<-read.csv(file="Data//crp_finals.csv", header=T)
#Rename the columns
colnames(meas_crpr) <- c("sample_code", "crp")
str(meas_crpr)
dim(meas_crpr)
#[1] 847   2

# merge dataframes
meas_d<-Reduce(function(x, y) merge(x, y, all=TRUE, by="sample_code"), 
               list(meas_mdacr_sp,meas_pest,meas_crpr))

str(meas_d)

#length(which(is.na(meas_d$creatinine)))
#14
#length(which(is.na(meas_d$crp)))
#94

# Delete observations 2126-2, 2216-2, 6086-2 because they are double 
# delete observations from participant 258 because excluded 
# delete observation 1004 because it is a non-existent sample
meas<-subset(meas_d,!((meas_d$sample_code == "2126-2")|
              (meas_d$sample_code == "2216-2")|
              (meas_d$sample_code == "6086-2")|
              (meas_d$sample_code == "2581")|
              (meas_d$sample_code == "2582")|
              (meas_d$sample_code == "2583")|
              (meas_d$sample_code == "1004")
              ))
str(meas)

# Extract 4 characters of sample code to have uniform codes 
# (transforms sample codes 2126-1 to 2126 sample_code needs to be a character) 
# rename to new variable to keep original
meas$sample_code0<-as.character(meas$sample_code) 
str(meas)
meas$sample_code<-substr(meas$sample_code0, start = 1, stop = 4)

# create sample number variable by deleting first three characters of sample_code
meas$sample_no_orig = substr(meas$sample_code, 4, 4)
str(meas)

# transform sample codes of participants who followed the opposite pattern 
# based on their school (203 & 259) so that they follow the same pattern with 
# the rest participants --- it only changes the last character, so that the 
# variable of phase can be created based on the conditional ifelse.
# These transformed sample codes are used only for the creation of the phase
# variable. For the statistical analysis, the original sample codes are used.
meas$sample_code1<-recode(meas$sample_code, "2032"="2034",
                          "2033"="2035", "2034"="2036",
                          "2035"="2032","2036"="2033",
                          "2593"="2595","2594"="2596",
                          "2595"="2592","2596"="2593")
str(meas)

# create ID variable by deleting last character of sample_code 
meas$Code<-as.character(meas$sample_code)
meas$Code = substr(meas$Code,1,nchar(meas$Code)-1)

# create sample number variable by deleting first three characters of sample_code1
meas$sample_no = substr(meas$sample_code1, 4, 4)

# create school variable by deleting last two characters of Code
meas$school = substr(meas$Code,1,nchar(meas$Code)-2) 

# create phase variable using conditions for school and sample no 
meas$phase<-ifelse(
  (meas$school==1 | meas$school==2 | meas$school==6) & 
    (meas$sample_no==1 | meas$sample_no==2 | meas$sample_no==3), 
  "Conventional", 
  ifelse((meas$school==3 | meas$school==4 | meas$school==5) & 
           (meas$sample_no==1 | meas$sample_no==5 | meas$sample_no==6),
         "Conventional", 
         "Organic"))

str(meas)

# change variable types into factors
meas[, c("Code", "sample_no", "sample_code","sample_code1","school","phase")]<-
  lapply(meas[, c("Code", "sample_no", "sample_code","sample_code1","school","phase")],
         factor)

str(meas)

# merge baseline data with measurement data
# use the set with the 149 participants for the merging 
meas1<-droplevels(merge(meas, mergedbaseline3, by="Code",all.y = TRUE))

# create subset for organic treatment only 
meas_org<-meas1 [which(meas1$phase == "Organic"),]
# create subset for conventional treatment only 
meas_con<-meas1 [which(meas1$phase == "Conventional"),]

```

Percentage of samples below the LOD and LOQ (Units Creatinine:g/l, 6CN:ug/l, 3PBA:ug/l, CRP:ng/L)

Below LOQ (only for the pesticides and only for the whole period)

6CN (<0.226), 3PBA (<0.146)

- 6CN: `r round(perc(meas1$cna, 0.226, dir = c("leq"), na.rm = TRUE), digits=2)`

- 3PBA: `r round(perc(meas1$pba, 0.146, dir = c("leq"), na.rm = TRUE), digits=2)`

Below LOD 

Creatinine (<0.25), 6CN (<0.075), 3PBA (<0.049), CRP (<62.5)

- Overall

+ Creatinine: `r round(perc(meas1$creatinine, 0.25, dir = c("leq"), na.rm = TRUE), digits=2)`

+ CRP: `r round(perc(meas1$crp, 62.5, dir = c("leq"), na.rm = TRUE), digits=2)`

+ 6CN: `r round(perc(meas1$cna, 0.075, dir = c("leq"), na.rm = TRUE), digits=2)`

+ 3PBA: `r round(perc(meas1$pba, 0.049, dir = c("leq"), na.rm = TRUE), digits=2)`

- Organic phase

+ CRP: `r round(perc(meas_org$crp, 62.5, dir = c("leq"), na.rm = TRUE), digits=2)`

+ 6CN: `r round(perc(meas_org$cna, 0.075, dir = c("leq"), na.rm = TRUE), digits=2)`

+ 3PBA: `r round(perc(meas_org$pba, 0.049, dir = c("leq"), na.rm = TRUE), digits=2)`

- Conventional phase (including the baseline)

+ CRP: `r round(perc(meas_con$crp, 62.5, dir = c("leq"), na.rm = TRUE), digits=2)`

+ 6CN: `r round(perc(meas_con$cna, 0.075, dir = c("leq"), na.rm = TRUE), digits=2)`

+ 3PBA: `r round(perc(meas_con$pba, 0.049, dir = c("leq"), na.rm = TRUE), digits=2)`


```{r 6CN values below LOD, include=FALSE}

rm(meas_con, meas_org) # remove the sets created before (no longer needed)

# since high percentage of samples below LOD for 6CN 
# check distribution of 6CN values in sampling points per group
# create categorical variables for below and above LOD values for 6-CN
# 6CN: <0.075: Below LOD, >0.075: Above LOD
meas1$cnlod <- cut(meas1$cna,
                  breaks=c(-Inf, 0.075, Inf),
                  labels=c("Below LOD","Above LOD"))

```

## Biomarkers descriptives

### Table 1 B - Percentage of 6-CN values below and above LOD by sample by group

`r kable(print(CreateTableOne("cnlod", data=meas1, strata=c("sample_no_orig","Group"), test=FALSE)), caption="Percentage of 6-CN values below and above LOD for 149 participants by group and by sample")`


```{r lod imputation, include=FALSE}

# only the <LOD are imputed

# replace values below LOD using ros function 
# for 3PBA (<0.049) because the percentage of values below LOD is higher than 20% and less than 70%.

# first create logical vector with TRUE indicating <LOD & FALSE >LOD
meas1$pba1<-ifelse(meas1$pba<0.049, TRUE, FALSE)

str(meas1$pba1)

# call the ros function
rospba<-ros(meas1$pba, meas1$pba1,forwardT="log", reverseT="exp", na.action=NULL)

# convert the modeled data in a ros model to a dataframe
rospba<-as.data.frame(rospba)

# sort meas1 dataframe by pba
meas1 <- meas1[order(meas1$pba),] 

# sort meas1 dataframe by pba
meas1 <- meas1[order(meas1$pba),] 

# merge meas1 with rospba
meas1<-cbind.data.frame(meas1,rospba)

# rename columns for rospba 
names(meas1)[names(meas1) == 'obs'] <- 'obs_pb'
names(meas1)[names(meas1) == 'censored'] <- 'censored_pb'
names(meas1)[names(meas1) == 'pp'] <- 'pp_pb'

# create variable for 3-pba with modelled values for below LOD 
# (values above LOD don't change in the model)
names(meas1)[names(meas1) == 'modeled'] <- 'pb'

#So, we will use the cr, cnlod and pb.

```

```{r biomarkers descriptives, include=FALSE}

#Rename the column name - creatinine to cr
colnames(meas1)[colnames(meas1) == "creatinine"] ="cr"

# From now on working on ng/L for 6CN, 3PBA
meas1$pb1000<-meas1$pb*1000
meas1$cn1000<-meas1$cna*1000


# adjust for creatinine 
meas1$adj_pb1000<-meas1$pb1000/meas1$cr
meas1$adj_cn1000<-meas1$cn1000/meas1$cr

meas1$adj_crp<-meas1$crp/meas1$cr

# select raw and adjusted variables
sel_raw<-c("pb1000","cn1000","crp","cr")

sel_adj<-c("adj_pb1000","adj_cn1000","adj_crp","cr")

# quantiles 

pb_q<-round(quantile(meas1$pb1000, c(0, .05,.25,.5,.75,.95,1),na.rm=TRUE), digits=2)
cn_q<-round(quantile(meas1$cn1000, c(0, .05,.25,.5,.75,.95,1),na.rm=TRUE), digits=2)

crp_q<-round(quantile(meas1$crp, c(0, .05,.25,.5,.75,.95,1), na.rm=TRUE), digits=2)
cr_q<-round(quantile(meas1$cr, c(0, .05,.25,.5,.75,.95,1), na.rm=TRUE), digits=2)

crpb_q<-round(quantile(meas1$adj_pb1000, c(0, .05,.25,.5,.75,.95,1),na.rm=TRUE), digits=2)
crcn_q<-round(quantile(meas1$adj_cn1000, c(0, .05,.25,.5,.75,.95,1),na.rm=TRUE), digits=2)

crp_q_adj<-round(quantile(meas1$adj_crp, c(0, .05,.25,.5,.75,.95,1),na.rm=TRUE), digits=2)

quantiles_raw<-rbind(pb_q,cn_q, crp_q, cr_q)
quantiles_adj<-rbind(crpb_q,crcn_q,crp_q_adj)

# quantiles by treatment 
measorg<-meas1 [which(meas1$phase == "Organic"),]

crp_qorg<-round(quantile(measorg$crp, c(0, .05,.25,.5,.75,.95,1), na.rm=TRUE), digits=2)

crp_qorg_adj<-round(quantile(measorg$adj_crp, c(0, .05,.25,.5,.75,.95,1), na.rm=TRUE), digits=2)

cr_qorg<-round(quantile(measorg$cr, c(0, .05,.25,.5,.75,.95,1), na.rm=TRUE), digits=2)

# select raw and adjusted quantiles in organic period
quantiles_rawo<-rbind(crp_qorg)
quantiles_adjo<-rbind(crp_qorg_adj)

meascon<-meas1 [which(meas1$phase == "Conventional"),]

crp_qcon<-round(quantile(meascon$crp, c(0, .05,.25,.5,.75,.95,1), na.rm=TRUE), digits=2)

crp_qcon_adj<-round(quantile(meascon$adj_crp, c(0, .05,.25,.5,.75,.95,1), na.rm=TRUE), digits=2)

cr_qcon<-round(quantile(meascon$cr, c(0, .05,.25,.5,.75,.95,1), na.rm=TRUE), digits=2)

# select raw and adjusted quantiles in conventional period
quantiles_rawc<-rbind(crp_qcon)
quantiles_adjc<-rbind(crp_qcon_adj)

```

### Table 2 - Percentiles of CRP, 6CN, 3PBA (overall)

- Note: the tables are based on the imputed data

- Units: CRP:ng/L, 3PBA:ng/L, 6-CN:ng/L

`r kable (print(quantiles_raw), caption="Quantiles of the raw concentrations")`

- Units: CRP:ng/L Cr, 3PBA:ng/L, 6-CN:ng/L

`r kable (print(quantiles_adj), caption="Quantiles of the creatinine adjusted concentrations")`

### Table 2 A - Percentiles of CRP during the organic treatment

- Note: the tables are based on the imputed data

- Units: CRP:ng/L

`r kable (print(quantiles_rawo), caption="Quantiles of the raw concentrations during the organic treatment")`

- Units: CRP:ng/L

`r kable (print(quantiles_adjo), caption="Quantiles of the creatinine adjusted concentrations during the organic treatment")`


### Table 2 B - Percentiles of CRP during the conventional treatment

- Note: the tables are based on the imputed data

- Units: CRP:ng/L

`r kable (print(quantiles_rawc), caption="Quantiles of the raw concentrations during the conventional treatment")`

- Units: CRP:ng/L Cr

`r kable (print(quantiles_adjc), caption="Quantiles of the creatinine adjusted concentrations during the conventional treatment")`


```{r Summary tables, include=FALSE}

# check normality of variables
qqnorm(meas1$pb1000) #NOT normal
qqnorm(meas1$crp) #NOT normal

qqnorm(meas1$adj_pb1000) #NOT normal
qqnorm(meas1$adj_crp) #NOT normal

# log-transform urinary data
meas1$logpb1000<-log(meas1$pb1000)
meas1$logcrp<-log(meas1$crp) 

# log-transform creatinine as per editor request
meas1$logcr<-log(meas1$cr)

meas1$logadj_pb1000<-log(meas1$adj_pb1000)
meas1$logadj_crp<-log(meas1$adj_crp) 

# check normality of variables again
qqnorm(meas1$logpb1000) # normal
hist(meas1$logpb1000)
qqnorm(meas1$logcrp)
hist(meas1$logcrp)

qqnorm(meas1$logadj_pb1000)
qqnorm(meas1$logadj_crp) # normal

hist(meas1$logadj_pb1000)
hist(meas1$logadj_crp)

```

### Table 3 - Summary tables of pesticide metabolites, creatinine and CRP - Overall

- Units: ng/L for CRP, 3-PBA, 6-CN and g/L for creatinine

`r kable(print(CreateContTable(sel_raw, data=meas1, testNonNormal = kruskal.test, smd=F), nonnormal=sel_raw), caption="Summary table of measurements - non-creatinine adjusted")`

- Units: ng/L for CRP, 3-PBA, 6-CN and g/L for creatinine

`r kable(print(CreateContTable(sel_adj, data=meas1, testNonNormal = kruskal.test, smd=F), nonnormal=sel_adj), caption="Summary table of measurements -Creatinine adjusted")`

### Table 3 A - Summary tables of pesticide metabolites, creatinine and CRP - By treatment

- Units: ng/L for CRP, 3-PBA, 6-CN and g/L for creatinine

`r kable(print(CreateContTable(sel_raw, strata=c("phase"),meas1, testNonNormal = kruskal.test, smd=F), nonnormal=sel_raw), caption="Summary table of measurements per phase - non-creatinine adjusted")`

- Units: ng/L for CRP, 3-PBA, 6-CN and g/L for creatinine

`r kable(print(CreateContTable(sel_adj, strata=c("phase"),meas1, testNonNormal = kruskal.test,smd=F), nonnormal=sel_adj), caption="Summary table of measurements per phase - Creatinine adjusted")`


## Boxplots

The boxplots are included in the script as additional material to what is included in the manuscript. The boxplots are not included in the manuscript.
In the boxplots of CRP the upper 5% of values have been excluded to make the plots easier to read. Imputed data are included. 

```{r boxplots, echo=FALSE}
# plot measurements per sample  per group
#length(which(is.na(meas1$crp)))
#7

#length(which(is.na(meas1$crp[meas1$crp < quantile(meas1$crp, 0.95,na.rm=TRUE)])))
#7

#is.numeric(meas1$crp)
#TRUE

#Remove the na of the meas$crp
meas_nonna<-meas1%>% filter(!is.na(crp))

ggplot(meas_nonna[meas_nonna$crp < quantile(meas_nonna$crp, 0.95,na.rm=TRUE), ], 
       aes(x=sample_no_orig, y=crp, fill=phase))+
  geom_boxplot()+
  facet_grid(~Group1)+
  theme_light()+
  ylab("crp (ng/l) - excl. >0.95")+
  xlab("")+
  scale_fill_brewer(palette ="Greens")+ 
  theme(legend.position="bottom")

meas_nonna_adj<-meas1%>% filter(!is.na(adj_crp))

ggplot(meas_nonna_adj[meas_nonna_adj$adj_crp < quantile(meas_nonna_adj$adj_crp, 0.95,na.rm=TRUE), ], 
       aes(x=sample_no_orig, y=adj_crp, fill=phase))+
  geom_boxplot()+
  facet_grid(~Group1)+
  theme_light()+
  ylab("creatinine adjusted crp - excl. >0.95")+
  xlab("")+
  scale_fill_brewer(palette ="Greens")+ 
  theme(legend.position="bottom")


ggplot(meas_nonna_adj[meas_nonna_adj$logadj_crp < quantile(meas_nonna_adj$logadj_crp, 0.95,na.rm=TRUE), ], 
       aes(x=sample_no_orig, y=logadj_crp, fill=phase))+
  geom_boxplot()+
  facet_grid(~Group1)+
  theme_light()+
  ylab("log creatinine adjusted crp- excl. >0.95")+
  xlab("")+
  scale_fill_brewer(palette ="Greens")+ 
  theme(legend.position="bottom")

```

```{r save, echo=FALSE}
dir_path<-getwd() # working dir to be used in making the new file bel
analysis_rds_dir<-paste0(dir_path, "/produced_data", sep="")
if(!dir.exists(analysis_rds_dir)) dir.create(analysis_rds_dir)

saveRDS(meas1, file.path(analysis_rds_dir, "meas1.rds"))
```

# Session information

```{r session}
sessionInfo()
```