---
title: "An anti-inflammatory response of an organic food intervention treatment by reducing pesticide exposures in children of Cyprus: a cluster-randomized crossover trial - Sensitivity Analysis Type 1 - part 1"
output:
  html_document: default
  word_document: default
authors: Cyprus International Institute for Environmental and Public Health, Cyprus
  University of Technology, Limassol, Cyprus.
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
  
## Sensitivity analysis Type 1 - part 1

Sensitivity analysis was carried out to assess if there was any influence of extreme values on the observed associations between exposure and outcome. CRP raw measurements higher than the 99th percentile were excluded from the analysis and all models were rerun. 

```{r libraries, include=FALSE}

## Prepare workspace & install libraries
rm(list = ls(all = TRUE))

ipak <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg)) 
    install.packages(new.pkg, dependencies = TRUE)
  sapply(pkg, require, character.only = TRUE)
}

# usage
packages <- c("readxl","plyr","dtplyr","data.table","tidyr","knitr",
              "tibble","Hmisc","tableone","compare","stargazer",
              "readr","ggplot2","sjPlot","tidyverse","lubridate",
              "eeptools", "Rcpp", "nlme", "lme4","lmerTest",
              "texreg","reshape","scales","FSA","NADA",
              "childsds", "sjPlot")

ipak(packages)

```

```{r baseline data reading and inspection, include=FALSE}
## Baseline questionnaire - Data inspection

#Three files named:
#- Baseline_Ques(Final) 250517-EA
#- Baseline_Ques(Final) 250517-SI
#- Baseline_Ques(Final) 250517-IC are located in the "Analysis" folder.

#The compiled data file "mergedbaseline" contains 191 observations and 
#72 variables. 

#Variable names, description, type and valid values are described in the file:
#"Variables characteristics" located in the same folder.

#The below actions have been done for data inspection:

#- Read data
#- Merge the 3 files into 1
#- Create school variable by deleting last 2 characters of Code
#- Change variable types into factors/characters
#- Format dates

#Additional actions:

#- Delete variable V20 because it contains no data 
#- Convert 555 (missing values) to NA
#- Calculate age at baseline

#+ first subset in order to omit NA
#+ merge baseline data with age data

#-Merge baseline data with samples per participant data. The file "Samples per participant.csv" 
#is located in the "Analysis" folder and contains 191 observations and 7 variables. Variable names, description, type and valid values are 
#described in the file "Variables characteristics" located in the same folder.

# read data & merge 
baseline1<-read.csv(file="Data//Baseline_Ques(Final) 250517-EA.csv", 
header=T, sep=";")
baseline2<-read.csv(file="Data//Baseline_Ques(Final) 250517-SI.csv", 
header=T, sep=";")
baseline3<-read.csv(file="Data//Baseline_Ques(Final) 250517-IC.csv", 
header=T, sep=";")

baseline<-rbind.data.frame(baseline1, baseline2, baseline3)

# merge baseline data with samples per participant data
samplesperpart<-read.csv(file="Data//Samples and days per participant.csv", header=T, sep=";")
# check variable types
str(samplesperpart)

# change code to factor
samplesperpart$Code<- as.factor(samplesperpart$Code)

mergedbaseline<-merge(baseline,samplesperpart, by="Code")

```

```{r recode variables, include=FALSE}
# create school variable by deleting last 2 characters of Code 
# Code needs to be a character
mergedbaseline$Code<-as.character(mergedbaseline$Code)
mergedbaseline$school<- substr(mergedbaseline$Code, 
                               start=1,
                               stop=nchar(mergedbaseline$Code)-2)


# change variable types into factors
mergedbaseline[, c("Editor", "Code", "Group", "Completion_study",
                   "medulev", "Other_educational_level_m",
                   "fedulev", "Gender", "handinmouth", "school")]<-lapply(
                     mergedbaseline[, c("Editor", "Code", "Group", 
                                        "Completion_study",
                                        "medulev", "Other_educational_level_m",
                                        "fedulev", "Gender", "handinmouth",
                                        "school")], factor)

# change variable types into characters
mergedbaseline[, c("city_village", "product_name",
                   "other_activity_name", "Other_Ctivity_name_1",
                   "other_sedent_name", "Other_sedent_name_1")]<- lapply(
                     mergedbaseline[, c("city_village",
                                        "product_name", "other_activity_name",
                                        "Other_Ctivity_name_1",
                                        "other_sedent_name", 
                                        "Other_sedent_name_1")], as.character)

# format dates 
mergedbaseline[, c("Qcompletion_date", "birthdate", "anthr_date_start",
                   "anthr_date_endorg","anthr_date_endconv")]<-lapply(
                     mergedbaseline[, c("Qcompletion_date", "birthdate",
                                        "anthr_date_start",
                                        "anthr_date_endorg",
                                        "anthr_date_endconv")],
                     as.Date, format="%d/%m/%Y")

# delete variable V20 because it contains no data
which(!is.na(mergedbaseline$V20))
#16  29 180 184 186 189 191)
mergedbaseline1<-subset(mergedbaseline, select=-V20) 

# convert 555 (missing values) to NA
mergedbaseline1[mergedbaseline1==555] <- NA
str(mergedbaseline1)

# create total samples number variable by adding all separate samples
mergedbaseline1$totsamples<-with(mergedbaseline1, S1 + S2 + S3 + S4 + S5 + S6)

# change totsamples variable  into factors
mergedbaseline1$totsamples<-as.factor(mergedbaseline1$totsamples)

# calculate age with decimals
# calculate the age at the start of the study
# subset with date start and birthdate in order to omit NA and calculate age_start
# Make NA this birthdate "1900-01-05"
mergedbaseline1[mergedbaseline1=="1900-01-05"] <- NA
baseline_age1<-subset(mergedbaseline1, 
                      select=c("Code","birthdate","anthr_date_start"))
baseline_age1 <- na.omit(baseline_age1)
baseline_age1$age_baseline<-with(baseline_age1, 
                                 eeptools::age_calc(dob=birthdate, 
                                          enddate=anthr_date_start, 
                                          units = "years"))

mergedbaseline_age1<-merge(mergedbaseline1, baseline_age1[c("Code","age_baseline")], by="Code",
                           all = TRUE)

# calculate the age at the end of the organic treatment
baseline_age2<-subset(mergedbaseline1, 
                      select=c("Code","birthdate","anthr_date_endorg"))
baseline_age2 <- na.omit(baseline_age2)
baseline_age2$age_endorg<-with(baseline_age2, 
                               eeptools::age_calc(dob=birthdate,
                                        enddate=anthr_date_endorg, 
                                        units = "years"))

mergedbaseline_age2<-merge(mergedbaseline_age1, 
                           baseline_age2[c("Code","age_endorg")], by="Code",all = TRUE)

# calculate the age at the end of the conventional treatment for Group 1 only
baseline_age3<-subset(mergedbaseline1,
                      select=c("Code","birthdate","anthr_date_endconv"))
baseline_age3 <- na.omit(baseline_age3)
baseline_age3$age_endcon<-with(baseline_age3, 
                               age_calc(dob=birthdate,
                                        enddate=anthr_date_endconv, 
                                        units = "years"))

mergedbaseline2<-merge(mergedbaseline_age2,
                       baseline_age3[c("Code","age_endcon")], by="Code",all = TRUE)

# revalue categorical variables to their actual values
mergedbaseline2$Sex <- revalue(mergedbaseline2$Gender, 
                               c("1"="Male", "2"="Female"))

# calculate BMI based on formula: weight/height^2
mergedbaseline2$BMI_start<-mergedbaseline2$Weight_start/
                          ((mergedbaseline2$Height_start/100)^2)
mergedbaseline2$BMI_endorg<-mergedbaseline2$Weight_endorg/
                          ((mergedbaseline2$Height_endorg/100)^2)
mergedbaseline2$BMI_endcon<-mergedbaseline2$Weight_endconv/
                          ((mergedbaseline2$Height_endconv/100)^2)

# calculate BMI SDS based on WHO 2007 growth references
mergedbaseline2$BMI_start_sds <- sds(mergedbaseline2$BMI_start,
                       age = mergedbaseline2$age_baseline,
                       sex = mergedbaseline2$Sex, 
                       male = "Male", female = "Female",
                       ref = who2007.ref,
                       item = "bmi",
                       type = "SDS")

mergedbaseline2$BMI_endorg_sds <- sds(mergedbaseline2$BMI_endorg,
                       age = mergedbaseline2$age_endorg,
                       sex = mergedbaseline2$Sex, 
                       male = "Male", female = "Female",
                       ref = who2007.ref,
                       item = "bmi",
                       type = "SDS")

mergedbaseline2$BMI_endcon_sds <- sds(mergedbaseline2$BMI_endcon,
                       age = mergedbaseline2$age_endcon,
                       sex = mergedbaseline2$Sex, 
                       male = "Male", female = "Female",
                       ref = who2007.ref,
                       item = "bmi",
                       type = "SDS")

# Check the new BMI variables
summary(mergedbaseline2$BMI_start_sds)
summary(mergedbaseline2$BMI_endorg_sds)
summary(mergedbaseline2$BMI_endcon_sds)

# create categories based on WHO cut-off points for 
# thinness, normal, overweight and obese.
# <-2: Thinness, 1>-2: Normal, >1: Overweight, >2: Obese
mergedbaseline2$BMI_for_age_start <- cut(mergedbaseline2$BMI_start_sds,
                               breaks=c(-Inf, -2, 1, 2, Inf),
                               labels=c("Thinness","Normal","Overweight","Obese"))
mergedbaseline2$BMI_for_age_endorg <- cut(mergedbaseline2$BMI_endorg_sds,
                               breaks=c(-Inf, -2, 1, 2, Inf),
                               labels=c("Thinness","Normal","Overweight","Obese"))
mergedbaseline2$BMI_for_age_endcon <- cut(mergedbaseline2$BMI_endcon_sds,
                               breaks=c(-Inf, -2, 1, 2, Inf),
                               labels=c("Thinness","Normal","Overweight","Obese"))

# revalue categorical variables to their actual values: 
# For 6=Other, 7=Missing and 8=NA use NA
mergedbaseline2$Group1 <- revalue(mergedbaseline2$Group, 
                                  c("1"="First organic", "2"="First conventional"))
mergedbaseline2$medulev1 <- revalue(mergedbaseline2$medulev, 
                                    c("1"="Primary", "2"="Secondary", 
                                      "3"="University/college", 
                                      "4"="Master/PhD", "6"= NA,"7"=NA, "8"=NA))
mergedbaseline2$fedulev1 <- revalue(mergedbaseline2$fedulev, 
                                    c("1"="Primary", "2"="Secondary", 
                                      "3"="University/college", 
                                      "4"="Master/PhD", "6"=NA,"7"=NA, "8"=NA))
mergedbaseline2$Handinmouth1 <- revalue(mergedbaseline2$handinmouth, 
                                        c("0"="No", "1"="Yes","7"=NA))

# recode Days_Org in categories 0-11 days (dropouts), 12-40 days (active participants)
mergedbaseline2$daysorg <- cut(mergedbaseline2$Days_Org,
                               breaks=c(-Inf, 11, 40),
                               labels=c("Dropouts","Active participants"))

# recode Days_Org in categories 0-11, 12-21, 22-28, 29-40
mergedbaseline2$daysorgcat <- cut(mergedbaseline2$Days_Org,
                               breaks=c(-Inf, 11, 21, 28, 40),
                               labels=c("0-11 days","12-21 days","22-28 days","29-40 days"))

# Last make the variable selection
selectionDescr<-c("Group1","medulev1","fedulev1","Sex","age_baseline",
                  "Waist_start", "Waist_endorg","Waist_endconv",
                  "BMI_start","BMI_endorg","BMI_endcon",
                  "BMI_for_age_start","BMI_for_age_endorg","BMI_for_age_endcon",
                  "totsamples","daysorgcat","Handinmouth1")

categorvariablesDescr<-c("Group1","medulev1","fedulev1","Sex",
                         "BMI_for_age_start","BMI_for_age_endorg","BMI_for_age_endcon",
                         "totsamples","daysorgcat","Handinmouth1")

# convert NA values to 0 in columns where the times per week are denoted in order to be included in the stats
mergedbaseline2[c("park_week","garden_week",
                  "run_week", "cyc_week", "bask_week", "voll_week", "swim_week",
                  "danc_week", "other_activity_time","Other_activity_time_1",
                  "tv_week", "comp_week",
                  "tablet_week", "mobile_week",
                  "other_sedent_week","Other_sedent_week_1")][is.na(mergedbaseline2[c("park_week","garden_week",
                  "run_week", "cyc_week", "bask_week", "voll_week", "swim_week",
                  "danc_week", "other_activity_time","Other_activity_time_1",
                  "tv_week", "comp_week",
                  "tablet_week", "mobile_week",
                  "other_sedent_week","Other_sedent_week_1")])] <- 0

# summarize time spent outside
mergedbaseline2$outside_time<-with(mergedbaseline2, park_week+garden_week)

# summarize time spent for physical activities
mergedbaseline2$physactiv_time<-with(mergedbaseline2, run_week+cyc_week+
                                       bask_week+voll_week+swim_week+danc_week+
                                       other_activity_time+Other_activity_time_1)

# summarize time spent for sedentary activities
mergedbaseline2$sedentactiv_time<-with(mergedbaseline2, tv_week+comp_week+
                                                           tablet_week+mobile_week+
                                                           other_sedent_week+
                                                           Other_sedent_week_1)

selectionActivities<-c("garden_week","park_week","run_week","cyc_week",
                       "bask_week","foot_week","voll_week","swim_week",
                       "danc_week","tv_week","comp_week", "tablet_week",
                       "mobile_week","outside_time","physactiv_time",
                       "sedentactiv_time")

# create subset with participants with more than 12 days in the organic phase (active)
mergedbaseline3<-droplevels(mergedbaseline2[which(mergedbaseline2$Days_Org > 11), ])

# check normality of variables
qqnorm(mergedbaseline3$Waist_start) 
hist(mergedbaseline3$Waist_start)

qqnorm(mergedbaseline3$physactiv_time)  
hist(mergedbaseline3$physactiv_time)

qqnorm(mergedbaseline3$sedentactiv_time) 
hist(mergedbaseline3$sedentactiv_time)

selectionnonnorm<-c("Waist_start", "Waist_endorg","Waist_endconv")
selectionactnonnorm<-c("physactiv_time","sedentactiv_time")

# create subset with participants with less than 12 days in the organic phase
mergedbaselinedropouts<-droplevels(mergedbaseline2[which(mergedbaseline2$Days_Org < 12), ])

# round age variables
mergedbaseline3$age_baseline<-round(mergedbaseline3$age_baseline)
mergedbaseline3$age_endorg<-round(mergedbaseline3$age_endorg)
mergedbaseline3$age_endcon<-round(mergedbaseline3$age_endcon)

```

## Urinary data

```{r Pesticides and biomarkers, include=FALSE}

rm(list=setdiff(ls(),c("mergedbaseline3", "mergedbaseline2")))

# Read measurements

meas_sa_mdacr<-read.csv(file="Data//mda_cr_meas.csv", header=T, sep=";")
str(meas_sa_mdacr)
dim(meas_sa_mdacr)
#941 3

#Keep the samples that the creatinine measurements are between 0.3 and 3 g/L.
meas_sa_mdacr_sp<-meas_sa_mdacr%>%
 filter(creatinine>0.29)%>%
  filter(creatinine<3.01)
dim(meas_sa_mdacr_sp)
#927 3
#There are 941-927=14 samples that the creatinine measurements are lower than 0.3 and upper than 3 g/L

meas_sa_pest<-read.csv(file="Data//Pesticides_meas.csv", header=T, sep=";")
str(meas_sa_pest)
dim(meas_sa_pest)
#941 3

#Units CRP:ng/L
meas_sa_crpr<-read.csv(file="Data//crp_finals.csv", header=T)
#Rename the columns
colnames(meas_sa_crpr) <- c("sample_code", "crp")
str(meas_sa_crpr)
dim(meas_sa_crpr)
#[1] 847   2

q99<-quantile(meas_sa_crpr$crp,0.99,na.rm=TRUE)

# Create the dataset for the sensitivity analysis
meas_sa_crpr_sa=meas_sa_crpr[which(meas_sa_crpr$crp<q99),]

# merge dataframes
meas_sa_d<-Reduce(function(x, y) merge(x, y, all=TRUE, by="sample_code"), 
               list(meas_sa_mdacr_sp,meas_sa_pest,meas_sa_crpr_sa))

str(meas_sa_d)

#length(which(is.na(meas_sa_d$creatinine)))
#14
#length(which(is.na(meas_sa_d$crp)))
#94

# Delete observations 2126-2, 2216-2, 6086-2 because they are double 
# observations from participant 258 because excluded & 
# delete observation 1004 because it is a non-existent sample
meas_sa<-subset(meas_sa_d,!((meas_sa_d$sample_code == "2126-2")|
              (meas_sa_d$sample_code == "2216-2")|
              (meas_sa_d$sample_code == "6086-2")|
              (meas_sa_d$sample_code == "2581")|
              (meas_sa_d$sample_code == "2582")|
              (meas_sa_d$sample_code == "2583")|
              (meas_sa_d$sample_code == "1004")
              ))
str(meas_sa)

# Extract 4 characters of sample code to have uniform codes 
# (transforms sample codes 2126-1 to 2126 sample_code needs to be a character) 
# rename to new variable to keep original
meas_sa$sample_code0<-as.character(meas_sa$sample_code) 
str(meas_sa)
meas_sa$sample_code<-substr(meas_sa$sample_code0, start = 1, stop = 4)

# create sample number variable by deleting first three characters of sample_code
meas_sa$sample_no_orig = substr(meas_sa$sample_code, 4, 4)
str(meas_sa)

# transform sample codes of participants who followed the opposite pattern 
# based on their school (203 & 259) so that they follow the same pattern with 
# the rest participants --- it only changes the last character, so that the 
# variable of phase can be created based on the conditional ifelse.
# These transformed sample codes are used only for the creation of the phase
# variable. For the statistical analysis, the original sample codes are used.
meas_sa$sample_code1<-recode(meas_sa$sample_code, "2032"="2034",
                          "2033"="2035", "2034"="2036",
                          "2035"="2032","2036"="2033",
                          "2593"="2595","2594"="2596",
                          "2595"="2592","2596"="2593")
str(meas_sa)

# create ID variable by deleting last character of sample_code 
meas_sa$Code<-as.character(meas_sa$sample_code)
meas_sa$Code = substr(meas_sa$Code,1,nchar(meas_sa$Code)-1)

# create sample number variable by deleting first three characters of sample_code1
meas_sa$sample_no = substr(meas_sa$sample_code1, 4, 4)

# create school variable by deleting last two characters of Code
meas_sa$school = substr(meas_sa$Code,1,nchar(meas_sa$Code)-2) 

# create phase variable using conditions for school and sample no 
meas_sa$phase<-ifelse(
  (meas_sa$school==1 | meas_sa$school==2 | meas_sa$school==6) & 
    (meas_sa$sample_no==1 | meas_sa$sample_no==2 | meas_sa$sample_no==3), 
  "Conventional", 
  ifelse((meas_sa$school==3 | meas_sa$school==4 | meas_sa$school==5) & 
           (meas_sa$sample_no==1 | meas_sa$sample_no==5 | meas_sa$sample_no==6),
         "Conventional", 
         "Organic"))

str(meas_sa)

# change variable types into factors
meas_sa[, c("Code", "sample_no", "sample_code","sample_code1","school","phase")]<-
  lapply(meas_sa[, c("Code", "sample_no", "sample_code","sample_code1","school","phase")],
         factor)

str(meas_sa)

# merge baseline data with meas_saurement data
# use the set with the 149 participants for the merging 
meas_sa1<-droplevels(merge(meas_sa, mergedbaseline3, by="Code",all.y = TRUE))

# create subset for organic treatment only 
meas_sa_org<-meas_sa1 [which(meas_sa1$phase == "Organic"),]
# create subset for conventional treatment only 
meas_sa_con<-meas_sa1 [which(meas_sa1$phase == "Conventional"),]

```

Percentage of samples below the LOD and LOQ (Units Creatinine:g/l, 6CN:ug/l, 3PBA:ug/l, CRP:ng/L)

Below LOQ (only for the pesticides and only for the whole period)

6CN (<0.226), 3PBA (<0.146)

- 6CN: `r round(perc(meas_sa1$cna, 0.226, dir = c("leq"), na.rm = TRUE), digits=2)`

- 3PBA: `r round(perc(meas_sa1$pba, 0.146, dir = c("leq"), na.rm = TRUE), digits=2)`

Below LOD 

Creatinine (<0.25), 6CN (<0.075), 3PBA (<0.049), CRP (<62.5)

- Overall

+ Creatinine: `r round(perc(meas_sa1$creatinine, 0.25, dir = c("leq"), na.rm = TRUE), digits=2)`

+ CRP: `r round(perc(meas_sa1$crp, 62.5, dir = c("leq"), na.rm = TRUE), digits=2)`

+ 6CN: `r round(perc(meas_sa1$cna, 0.075, dir = c("leq"), na.rm = TRUE), digits=2)`

+ 3PBA: `r round(perc(meas_sa1$pba, 0.049, dir = c("leq"), na.rm = TRUE), digits=2)`

- Organic phase

+ CRP: `r round(perc(meas_sa_org$crp, 62.5, dir = c("leq"), na.rm = TRUE), digits=2)`

+ 6CN: `r round(perc(meas_sa_org$cna, 0.075, dir = c("leq"), na.rm = TRUE), digits=2)`

+ 3PBA: `r round(perc(meas_sa_org$pba, 0.049, dir = c("leq"), na.rm = TRUE), digits=2)`

- Conventional phase (including the baseline)

+ CRP: `r round(perc(meas_sa_con$crp, 62.5, dir = c("leq"), na.rm = TRUE), digits=2)`

+ 6CN: `r round(perc(meas_sa_con$cna, 0.075, dir = c("leq"), na.rm = TRUE), digits=2)`

+ 3PBA: `r round(perc(meas_sa_con$pba, 0.049, dir = c("leq"), na.rm = TRUE), digits=2)`


```{r 6CN values below LOD, include=FALSE}

rm(meas_sa_con, meas_sa_org) # remove the sets created before (no longer needed)

# since high percentage of samples below LOD for 6CN 
# check distribution of 6CN values in sampling points per group
# create categorical variables for below and above LOD values for 6-CN
# 6CN: <0.075: Below LOD, >0.075: Above LOD
meas_sa1$cnlod <- cut(meas_sa1$cna,
                  breaks=c(-Inf, 0.075, Inf),
                  labels=c("Below LOD","Above LOD"))

```

## Biomarkers descriptives

### Table 1 - Percentage of 6-CN values below and above LOD by sample by group

`r kable(print(CreateTableOne("cnlod", data=meas_sa1, strata=c("sample_no_orig","Group"), test=FALSE)), caption="Percentage of 6-CN values below and above LOD for 149 participants by group and by sample")`


```{r lod imputation, include=FALSE}

# only the <LOD imputed

# replace values below LOD with LOD/2 for: 
# Creatinine (<0.25 = 0.13) 
# since the percentage of values below LOD is about 1%
meas_sa1<-meas_sa1%>%
  mutate(cr=ifelse(creatinine > 0.25,creatinine,0.13))

# replace values below LOD imputation using ros function 
# for 6CN (<0.075), 3PBA (<0.049) because the percentage of values below LOD is higher than 20%

# first create logical vector with TRUE indicating <LOD & FALSE >LOD
meas_sa1$pba1<-ifelse(meas_sa1$pba<0.049, TRUE, FALSE)

str(meas_sa1$pba1)

# call the ros function
rospba<-ros(meas_sa1$pba, meas_sa1$pba1,forwardT="log", reverseT="exp", na.action=NULL)

# convert the modeled data in a ros model to a dataframe
rospba<-as.data.frame(rospba)

# sort meas_sa1 dataframe by pba
meas_sa1 <- meas_sa1[order(meas_sa1$pba),] 

# sort meas_sa1 dataframe by pba
meas_sa1 <- meas_sa1[order(meas_sa1$pba),] 

# merge meas_sa1 with rospba
meas_sa1<-cbind.data.frame(meas_sa1,rospba)

# rename columns for rospba 
names(meas_sa1)[names(meas_sa1) == 'obs'] <- 'obs_pb'
names(meas_sa1)[names(meas_sa1) == 'censored'] <- 'censored_pb'
names(meas_sa1)[names(meas_sa1) == 'pp'] <- 'pp_pb'

# create variable for 3-pba with modelled values for below LOD 
# (values above LOD don't change in the model)
names(meas_sa1)[names(meas_sa1) == 'modeled'] <- 'pb'

#So, we will use the cr, cnlod and pb.

```

```{r biomarkers descriptives, include=FALSE}

# From now on working on ng/L for 6CN, 3PBA
meas_sa1$pb1000<-meas_sa1$pb*1000

# adjust for creatinine 
meas_sa1$adj_pb1000<-meas_sa1$pb1000/meas_sa1$cr
meas_sa1$adj_crp<-meas_sa1$crp/meas_sa1$cr

# select raw and adjusted variables
sel_raw<-c("pb1000","crp","cr")

sel_adj<-c("adj_pb1000","adj_crp","cr")

# quantiles - removed creatinine and rearranged them to follow the necessary table order

pb_q<-round(quantile(meas_sa1$pb1000, c(0, .05,.25,.5,.75,.95,1),na.rm=TRUE), digits=2)
crp_q<-round(quantile(meas_sa1$crp, c(0, .05,.25,.5,.75,.95,1), na.rm=TRUE), digits=2)
cr_q<-round(quantile(meas_sa1$cr, c(0, .05,.25,.5,.75,.95,1), na.rm=TRUE), digits=2)

crpb_q<-round(quantile(meas_sa1$adj_pb1000, c(0, .05,.25,.5,.75,.95,1),na.rm=TRUE), digits=2)
crp_q_adj<-round(quantile(meas_sa1$adj_crp, c(0, .05,.25,.5,.75,.95,1),na.rm=TRUE), digits=2)

quantiles_raw<-rbind(pb_q,crp_q, cr_q)
quantiles_adj<-rbind(crpb_q,crp_q_adj)

# quantiles by treatment 
meas_saorg<-meas_sa1 [which(meas_sa1$phase == "Organic"),]

crp_qorg<-round(quantile(meas_saorg$crp, c(0, .05,.25,.5,.75,.95,1), na.rm=TRUE), digits=2)

crp_qorg_adj<-round(quantile(meas_saorg$adj_crp, c(0, .05,.25,.5,.75,.95,1), na.rm=TRUE), digits=2)

cr_qorg<-round(quantile(meas_saorg$cr, c(0, .05,.25,.5,.75,.95,1), na.rm=TRUE), digits=2)

# select raw and adjusted quantiles in organic period
quantiles_rawo<-rbind(crp_qorg)
quantiles_adjo<-rbind(crp_qorg_adj)

meas_sacon<-meas_sa1 [which(meas_sa1$phase == "Conventional"),]

crp_qcon<-round(quantile(meas_sacon$crp, c(0, .05,.25,.5,.75,.95,1), na.rm=TRUE), digits=2)

crp_qcon_adj<-round(quantile(meas_sacon$adj_crp, c(0, .05,.25,.5,.75,.95,1), na.rm=TRUE), digits=2)

cr_qcon<-round(quantile(meas_sacon$cr, c(0, .05,.25,.5,.75,.95,1), na.rm=TRUE), digits=2)

# select raw and adjusted quantiles in conventional period
quantiles_rawc<-rbind(crp_qcon)
quantiles_adjc<-rbind(crp_qcon_adj)

```

### Table 2 - Percentiles of CRP, 6CN, 3PBA (overall)

- Note: the tables are based on the imputed data

- Units: CRP:ng/L, 3PBA:ng/L

`r kable (print(quantiles_raw), caption="Quantiles of the raw concentrations")`

- Units: crp:ng/L Cr, 3PBA:ng/L

`r kable (print(quantiles_adj), caption="Quantiles of the creatinine adjusted concentrations")`

### Table 2 A - Percentiles of CRP during the organic treatment

- Note: the tables are based on the imputed data

- Units: CRP:ng/L

`r kable (print(quantiles_rawo), caption="Quantiles of the raw concentrations during the organic treatment")`

- Units: CRP:ng/L

`r kable (print(quantiles_adjo), caption="Quantiles of the creatinine adjusted concentrations during the organic treatment")`


### Table 2 B - Percentiles of CRP during the conventional treatment

- Note: the tables are based on the imputed data

- Units: crp:ng/L

`r kable (print(quantiles_rawc), caption="Quantiles of the raw concentrations during the conventional treatment")`

- Units: crp:ng/L Cr

`r kable (print(quantiles_adjc), caption="Quantiles of the creatinine adjusted concentrations during the conventional treatment")`


```{r Summary tables, include=FALSE}

# For % change between baseline and after 40 days of organic diet, 
# we need the sample before the 
# beginning of the organic treatment and the last sample of the organic treatment

# sample before organic treatment for Group 1 -> baseline or sample no=1
# last sample of organic treatment for Group 1 -> sample no=4
# sample before organic treatment for Group 2 -> last sample of conventional treatment or 
# sample no=3
# last sample of organic treatment for Group 2 -> sample no=6

# check normality of variables
qqnorm(meas_sa1$pb1000)#NOT normal
qqnorm(meas_sa1$crp) #NOT normal

qqnorm(meas_sa1$adj_pb1000)#NOT normal
qqnorm(meas_sa1$adj_crp)#NOT normal

# log-transform urinary data
meas_sa1$logpb1000<-log(meas_sa1$pb1000)
meas_sa1$logcrp<-log(meas_sa1$crp) 

meas_sa1$logadj_pb1000<-log(meas_sa1$adj_pb1000)
meas_sa1$logadj_crp<-log(meas_sa1$adj_crp) 

# check normality of variables again
qqnorm(meas_sa1$logpb1000)# normal
hist(meas_sa1$logpb1000)
qqnorm(meas_sa1$logcrp)
hist(meas_sa1$logcrp)

qqnorm(meas_sa1$logadj_pb1000)
qqnorm(meas_sa1$logadj_crp)# normal

hist(meas_sa1$logadj_pb1000)
hist(meas_sa1$logadj_crp)

```

### Table 3 - Summary tables of pesticide metabolites, creatinine and CRP

- Note: the tables are based on the imputed data

- Units: ng/L for CRP, 3-PBA, g/L for creatinine

`r kable(print(CreateContTable(sel_raw, data=meas_sa1, testNonNormal = kruskal.test, smd=F), nonnormal=sel_raw), caption="Summary table of meas_saurements - non-creatinine adjusted")`

- Units: ng/L for CRP, 3-PBA, g/L for creatinine

`r kable(print(CreateContTable(sel_adj, data=meas_sa1, testNonNormal = kruskal.test, smd=F), nonnormal=sel_raw), caption="Summary table of meas_saurements -Creatinine adjusted")`

### Table 3 A - Summary tables of pesticide metabolites, creatinine and CRP by treatment

- Note: the tables are based on the imputed data

- Units: ng/L for CRP, 3-PBA, g/L for creatinine

`r kable(print(CreateContTable(sel_raw, strata=c("phase"),meas_sa1, testNonNormal = kruskal.test, smd=F), nonnormal=sel_raw), caption="Summary table of meas_saurements per phase - non-creatinine adjusted")`

- Units: ng/L for CRP, 3-PBA, g/L for creatinine

`r kable(print(CreateContTable(sel_adj, strata=c("phase"),meas_sa1, testNonNormal = kruskal.test,smd=F), nonnormal=sel_adj), caption="Summary table of meas_saurements per phase - Creatinine adjusted")`


## Boxplots

The boxplots are included in the script as additional material to what is included in the manuscript. The boxplots are not included in the manuscript.
In the boxplots of CRP the upper 5% of values have been excluded to make the plots easier to read. Imputed data are included. 

```{r boxplots, echo=FALSE}
# plot meas_saurements per sample  per group
length(which(is.na(meas_sa1$crp)))
#7

length(which(is.na(meas_sa1$crp[meas_sa1$crp < quantile(meas_sa1$crp, 0.95,na.rm=TRUE)])))
#7

is.numeric(meas_sa1$crp)
#TRUE

#Remove the na of the meas_sa$crp

meas_sa_nonna<-meas_sa1%>% filter(!is.na(crp))

ggplot(meas_sa_nonna[meas_sa_nonna$crp < quantile(meas_sa_nonna$crp, 0.95,na.rm=TRUE), ], 
       aes(x=sample_no_orig, y=crp, fill=phase))+
  geom_boxplot()+
  facet_grid(~Group1)+
  theme_light()+
  ylab("crp (ng/l) - excl. >0.95")+
  xlab("")+
  scale_fill_brewer(palette ="Greens")+ 
  theme(legend.position="bottom")

meas_sa_nonna_adj<-meas_sa1%>% filter(!is.na(adj_crp))

ggplot(meas_sa_nonna_adj[meas_sa_nonna_adj$adj_crp < quantile(meas_sa_nonna_adj$adj_crp, 0.95,na.rm=TRUE), ], 
       aes(x=sample_no_orig, y=adj_crp, fill=phase))+
  geom_boxplot()+
  facet_grid(~Group1)+
  theme_light()+
  ylab("creatinine adjusted crp - excl. >0.95")+
  xlab("")+
  scale_fill_brewer(palette ="Greens")+ 
  theme(legend.position="bottom")


ggplot(meas_sa_nonna_adj[meas_sa_nonna_adj$logadj_crp < quantile(meas_sa_nonna_adj$logadj_crp, 0.95,na.rm=TRUE), ], 
       aes(x=sample_no_orig, y=logadj_crp, fill=phase))+
  geom_boxplot()+
  facet_grid(~Group1)+
  theme_light()+
  ylab("log creatinine adjusted crp- excl. >0.95")+
  xlab("")+
  scale_fill_brewer(palette ="Greens")+ 
  theme(legend.position="bottom")

```

```{r save}
dir_path<-getwd() # working dir to be used in making the new file bel
analysis_rds_dir<-paste0(dir_path, "/produced_data", sep="")
if(!dir.exists(analysis_rds_dir)) dir.create(analysis_rds_dir)

saveRDS(meas_sa1, file.path(analysis_rds_dir, "meas_sa1.rds"))
```

# Session information

```{r session}
sessionInfo()
```