---
title: "An anti-inflammatory response of an organic food intervention treatment by reducing pesticide exposures in children of Cyprus: a cluster-randomized crossover trial - Sensitivity Analysis Type 1 - part 2" 
output:
  html_document:
    toc: yes
    toc_float: yes
  word_document:
    toc: yes
authors: Cyprus International Institute for Environmental and Public Health, Cyprus
  University of Technology, Limassol, Cyprus.
---

# Sensitivity analysis Type 1 - Regression analysis (mixed effects models)

This script includes the sensitivity analysis type 1-regression analysis (Part 2).
The output of the Part 1 (sensitivity analysis type 1 - part 1) script needs to be created (produced_data/meas_sa1).

```{r libraries, include=FALSE}

## Prepare workspace & install libraries

rm(list = ls(all = TRUE))

ipak <- function(pkg) {
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg)) {
    install.packages(new.pkg, dependencies = TRUE)
  }
  sapply(pkg, require, character.only = TRUE)
}

# usage
packages <- c(
  "readxl", "plyr", "dtplyr", "data.table", "tidyr", "knitr",
  "tibble", "Hmisc", "tableone", "compare", "stargazer",
  "readr", "ggplot2", "sjPlot", "tidyverse", "lubridate",
  "eeptools", "Rcpp", "nlme", "lme4", "lmerTest",
  "texreg", "reshape", "scales", "FSA", "NADA",
  "childsds", "sjPlot","ggpubr"
)

ipak(packages)

```

```{r linear mixed effect models - prepare database, echo=FALSE, include=FALSE}
# read the output of part1
dir_path <- getwd()
meas_sa1 <- readRDS(paste0(dir_path, "/produced_data/meas_sa1.rds", sep = ""))

# create dataset with variables needed for the regression analysis (we keep the log transformed crp and pesticides)
lmmdata <- subset(meas_sa1, select = c(
  "Code", "phase", "school.x",
  "sample_no_orig", "Group",
  "age_baseline", "age_endcon", "age_endorg",
  "Sex","logadj_crp","logcrp","cnlod","adj_pb1000","pb1000","cr" 
))

# scale and center continuous variables
lmmdata$logadj_crp_S <- scale(lmmdata$logadj_crp)

lmmdata$logcrp_S <- scale(lmmdata$logcrp)

lmmdata$age_baseline_S <- scale(lmmdata$age_baseline)

lmmdata$adj_pb1000_S<-scale(lmmdata$adj_pb1000)

lmmdata$pb1000_S<-scale(lmmdata$pb1000)

lmmdata$creatinine_S<-scale(lmmdata$cr)

# create a variable that denotes the "before(=1)" sample (which is sample no.1)
# and the "after(=2)" sample (no.4 for Group 1 and no.6 for Group 2)
# all other samples set to '9'
# This variable is used for the subsetting of the BMI dataset
lmmdata$before_after <- ifelse((lmmdata$sample_no_orig == 1),
  1,
  ifelse((lmmdata$sample_no_orig == 4 &
    lmmdata$Group == "1"),
  2,
  ifelse((lmmdata$sample_no_orig == 6 &
    lmmdata$Group == "2"),
  2,
  9
  )
  )
)

# reform dataset - Change variable types into factors/numeric
lmmdata <- droplevels(lmmdata)
lmmdata$ID <- as.factor(paste("P", lmmdata$Code, sep = ""))
lmmdata$SCID <- as.factor(paste("S", lmmdata$school.x, sep = ""))
lmmdata$GID <- as.factor(paste("G", lmmdata$Group, sep = ""))
lmmdata$ST <- as.numeric(lmmdata$sample_no_orig)

# import sample dates
#- from the original file
#- Kept only white cap rows
#- Deleted all columns besides code and date
#- 941 rows with participant samples left
#- Deleted double samples 6086-2, 2126-2, 2216-2
#- calculate days from baseline
#- Calculate days from treatment only for conventional phase

# read excel file
sample_dates <- read_excel("Data\\sample_dates.xlsx",
  col_types = c("text", "date")
)

# add ID
sample_dates$ID <- paste("P", substr(sample_dates$sample_code,
  start = 1,
  stop = 3
), sep = "")
# add sample number
sample_dates$sample_no_orig <- as.character(substr(sample_dates$sample_code,
  start = 4, stop = 4
))
# create new dataset with ID, sample number and sampling date
unique_sample_dates <- unique(sample_dates[, c(
  "ID", "sample_no_orig",
  "sampling_date"
)])
# create dataset that has 1 new variable:
# Diff: the difference in days between each sample for each participant
df <- unique_sample_dates %>%
  group_by(ID) %>%
  mutate(Diff = c(0, diff(sampling_date)))

# merge df dataset with lmmdata dataset by ID and sample number
lmmdata_dates <- inner_join(df, lmmdata, by = c("ID", "sample_no_orig"))

# subset and include only conventional treatment samples
lmmdata_dates1 <- subset(lmmdata_dates, lmmdata_dates$phase == "Conventional")

# Using the variable for the number of days (Diff) to sum the cummulative
# number of days from the start of the conventional treatment by participant
lmmdata_dates2 <- lmmdata_dates1 %>%
  group_by(ID) %>%
  mutate(DaysfromTreatment = cumsum(Diff))

# summary(lmmdata_dates2$DaysfromTreatment)

# remove sets no longer needed
rm(lmmdata_dates, lmmdata_dates1)

# For some participants the organic phase started a few days after the collection of the
# baseline sample or the last conventional phase sample. The date of the start of the
# organic phase has been used as a reference to calculate the number of days from the
# start of organic phase. The corrected dates (differing in the sample 1 - baseline -,
# or sample 3 - the last sample of the conventional phase -) are included in this
# dataset (referred to as "organic treatment start dates")

# import organic treatment start dates
# sampling date different than start of organic treatment so for actual days of organic treatment this file was used
# calculate days from baseline
# Calculate days from treatment only for organic phase

# read excel file
org_dates <- read_excel("Data\\org_start_dates.xlsx",
  col_types = c("text", "date")
)

# add ID
org_dates$ID <- paste("P", substr(org_dates$sample_code,
  start = 1,
  stop = 3
), sep = "")
# add sample no
org_dates$sample_no_orig <- as.character(substr(org_dates$sample_code,
  start = 4, stop = 4
))
# create new dataset with ID, sample number and sampling date
unique_org_dates <- unique(org_dates[, c(
  "ID", "sample_no_orig",
  "sampling_date"
)])
# create dataset that has 1 new variable:
# Diff: the difference in days between each sampling for each participant
df1 <- unique_org_dates %>%
  group_by(ID) %>%
  mutate(Diff = c(0, diff(sampling_date)))

# merge df dataset with lmmdata dataset by ID and sample number
lmmdata_dates3 <- inner_join(df1, lmmdata, by = c("ID", "sample_no_orig"))

# subset and include only organic treatment samples
lmmdata_dates4 <- subset(lmmdata_dates3, lmmdata_dates3$phase == "Organic")

# Using the variable for the number of days (Diff) to sum the cummulative
# number of days from the start of the organic treatment by participant
lmmdata_dates5 <- lmmdata_dates4 %>%
  group_by(ID) %>%
  mutate(DaysfromTreatment = cumsum(Diff))

# summary(lmmdata_dates5$DaysfromTreatment)

# remove datasets not needed
rm(lmmdata_dates3, lmmdata_dates4)

# merge datasets to include all samples
lmmdata_dates_final <- rbind.data.frame(lmmdata_dates2, lmmdata_dates5)

# remove datasets not needed
rm(lmmdata_dates2, lmmdata_dates5)

# exclude baseline sample
lmmdata_noBL <- subset(lmmdata_dates_final, ST >= 2)

# include baseline sample and biomarkers
lmmdata_onlyBL <- subset(
  lmmdata_dates_final[c(
    "ID", "logadj_crp_S", "ST","logcrp_S"
  )],
  ST <= 1
)

# str(lmmdata_onlyBL)

# merge datasets without baseline and with baseline for biomarkers
lmmdata_dates_MOD <- merge(lmmdata_onlyBL, lmmdata_noBL, by = "ID", all.y = TRUE, suffixes = c("_BL", "_TR"))
# the baseline_samples end in _BL

# check the variable DaysfromTreatment
summary(lmmdata_dates_MOD$DaysfromTreatment)

# remove sets no longer needed and keep only the ones that are going to be used
rm(list = setdiff(
  ls(),
  c(
    "lmmdata_dates_MOD", # dataset that includes baseline as a variable
    "lmmdata_noBL", # baseline excluded
    "lmmdata_dates_final" # all data
  )
))

# vector with the titles of the rows used in the model specifications
rowtitles <- c(
  "Number of measurements",
  "Number of participants",
  "Number of schools"
)

# Function to exponentiate the estimates and CIs of the phase
OrganicPhaseExpEST<-function(x) {
  est<-subset(x, Row.names=="phaseOrganic", select=c("Estimate","CI2.5","CI97.5"))
  est2<-round(exp(est), digits=3)
  rm(est)
  names(est2)<-c("GMR", "CI2.5","CI97.5")
  est2
}

# Function to exponentiate the estimates and CIs of the phase
logadj_crp_S_TRExpEST<-function(x) {
  est<-subset(x, Row.names=="logadj_crp_S_TR", select=c("Estimate","CI2.5","CI97.5"))
  est2<-round(exp(est), digits=3)
  rm(est)
  names(est2)<-c("GMR", "CI2.5","CI97.5")
  est2
}

#Create a new dataframe the spesific column of lmmdata_dates_MOD and keep the non na values
visua<-lmmdata_dates_MOD%>% select(ID,logadj_crp_S_TR, logcrp_S_TR,phase,DaysfromTreatment)%>%filter(!is.na(logadj_crp_S_TR))

```
## Visualization 

Visualization the association between the creatinine adjusted crp measurements and the time and the non creatinine crp adjusted measurements and the days from treatment.

*  Both measurements (creatinine adjusted crp and non creatinine adjusted crp) are log transformed.

__Plot 1. Association between the CRP measurements (adjusted and non adjusted) and the days from treatment for  the organic phase__ 

```{r visua_crp organic period, echo=FALSE}

visua_org<-visua%>%filter(phase=="Organic")

org1<-ggplot(data = visua_org, aes(x = DaysfromTreatment, y = logadj_crp_S_TR))+ geom_point(color="darkred")+labs(x = "Days from treatment",y = "Creatinine-adj, scaled and centered CRP measurements") 

org2<-ggplot(data = visua_org, aes(x = DaysfromTreatment, y = logcrp_S_TR))+ geom_point(color="darkblue")+labs(x = "Days from treatment",y = "Non-creatinine adj, scaled and centered CRP measurements") 

plot_org<-ggarrange(org1, org2,ncol = 2, nrow = 1)

annotate_figure(plot_org, top = text_grob("Organic phase", color = "black", face = "bold", size = 14))

```
__Plot 2. Association between the CRP measurements (adjusted and non adjusted) and the days from treatment for the conventional phase__

```{r visua_crp conventional period, echo=FALSE}

visua_con<-visua%>%filter(phase=="Conventional")

conv1<-ggplot(data = visua_con, aes(x = DaysfromTreatment, y = logadj_crp_S_TR))+ geom_point(color="darkred")+labs(x = "Days from treatment",y = "Creatinine-adj, scaled and centered CRP measurements") 

conv2<-ggplot(data = visua_con, aes(x = DaysfromTreatment, y = logcrp_S_TR))+ geom_point(color="darkblue")+labs(x = "Days from treatments",y = "Non-creatinine adj, scaled and centered CRP measurements") 

plot_con<-ggarrange(conv1, conv2,ncol = 2, nrow = 1)

annotate_figure(plot_con, top = text_grob("Conventional phase", color = "black", face = "bold", size = 14))

```

__Plot 3. Association between the creatinine adjusted CRP measurements and the days from treatment for each phase__

```{r plot, echo=FALSE}

ggplot(data = visua, aes(x = DaysfromTreatment, y = logadj_crp_S_TR, color = phase))+ geom_point()+facet_wrap(facets = vars(phase))+labs(x = "Days from treatment",y = "Creatinine-adj, scaled and centered CRP measurements") 

```

__Plot 4. Association between the non-creatinine adjusted CRP measurements and the days from treatment for each phase__

``` {r plot 2, echo=FALSE}

ggplot(data = visua, aes(x = DaysfromTreatment, y = logcrp_S_TR, color = phase))+ geom_point()+facet_wrap(facets = vars(phase))+labs(x = "Days from treatment",y = "Non-creatinine adj, scaled and centered CRP measurements") 

```

## Sensitivity analysis Type 1 - Models details

As the main analysis:

1. All models were adjusted for the baseline value of the outcome.

2. All models were adjusted for the creatinine levels using two ways:

*  CRP and pesticides measurements were divided with the respective creatinine measurements
*  Creatinine was included as a fixed effect

Initial model (model for objective 1):

 + Random part: participants
 + Fixed effect: treatment condition (organic or conventional) and time (days of treatment, where time=0 was the start of the treatment)
 + Interaction term: time*phase
*  Check interaction term:
  + if p-value>0.05 remove from the model
  
### Log creatinine adjusted CRP ~ baseline log creatinine adjusted CRP + DaysfromTreatment + phase + (phase*DaysfromTreatment if p<0.05) + random(...)

__Initial model (including the interaction term)__

```{r  crpmodel_INT, echo=FALSE, include=FALSE}

# model with interaction term
model_crp <- lmer(logadj_crp_S_TR ~ logadj_crp_S_BL + DaysfromTreatment + phase + DaysfromTreatment*phase + (1 | ID), lmmdata_dates_MOD, na.action = na.exclude)

# summary(model_crp)
coefs_crp_INT <- merge(
  as.data.frame(
    round(
      # getting from the model coefficients table
      # the fixed effects names --- row.names
      # the estimate --- col. 1
      # the std. error --- col. 2
      # the p-value --- col. 5
      coef(summary(model_crp))[, c(1, 2, 5)],
      digits = 3
    )
  ),
  as.data.frame(
    round(
      # getting the confidence intervals omitting the first 3 rows that
      # refer to sigma (it is NA in the Wald method)
      confint.merMod(model_crp, method = "Wald"),
      digits = 3
    )[-c(1:3), ]
  ),
  by = "row.names"
)

# change the names in the columns of the CIs
setnames(coefs_crp_INT, old = c("Pr(>|t|)", "2.5 %", "97.5 %"), new = c("p", "CI2.5", "CI97.5"))

coefs_crp_INT$Model <- "crpadjusted_INT"

# revalue the fixed effects names
coefs_crp_INT$FixedEffects <- revalue(
  coefs_crp_INT$Row.names,
  c(
    "(Intercept)" = "Intercept",
    "DaysfromTreatment" = "Time (days)",
    "DaysfromTreatment:phaseOrganic"="Days*Phase",
    "phaseOrganic" = "Organic diet",
    "logadj_crp_S_BL" = "Baseline crp levels"
  )
)

model_specs_crp <- c(
  # Number of measurements taken from the length of the outcome variable used in the model
  length(model_crp@frame$logadj_crp_S_TR),
  # Number of participants taken from the levels of the ID variables
  nlevels(model_crp@frame$ID),
  # Number of schools taken from the model details
  nlevels(model_crp@flist$SCID)
)

# add the row titles to the model specs
obs_dets_crp <- data.frame(rowtitles, model_specs_crp)
names(obs_dets_crp) <- c("Model details", "")

```

```{r tabmpdel Model 1, echo=FALSE, include=TRUE}
# crp
tab_model(model_crp,
  show.icc = TRUE,
  digits = 3, show.obs = TRUE, show.re.var = TRUE,
  title = paste(model_crp@call$formula[3]))
# model has been inspected: interaction term p-value<0.05

```

```{r model_crp, echo=FALSE}

plot(model_crp, resid(.) ~ fitted(.) | phase)

plot(model_crp, resid(.) ~ DaysfromTreatment)

qqnorm(resid(model_crp))

```

### Log CRP ~ baseline Log CRP + DaysfromTreatment + phase + creatinine_S + (phase*DaysfromTreatment if p<0.05) + random(...)

```{r  crpmodel_INT logcr int, echo=FALSE, include=FALSE}

# model with interaction term
model_crp_cr_INT <- lmer(logcrp_S_TR ~ logcrp_S_BL + DaysfromTreatment + phase + creatinine_S + phase*DaysfromTreatment +(1 | ID), lmmdata_dates_MOD, na.action = na.exclude)

```

```{r tabmpdel Model logcr int, echo=FALSE, include=TRUE}
# crp

#tab_model(model_crp_cr_INT,
#  show.icc = TRUE,
#  digits = 3, show.obs = TRUE, show.re.var = TRUE,
#  title = paste(model_crp_cr_INT@call$formula[3]))
# model has been inspected: interaction term p-value>0.05

```
```{r model_crp_logcr_INT, echo=FALSE}

#plot(model_crp_cr_INT, resid(.) ~ fitted(.) | phase)

#plot(model_crp_cr_INT, resid(.) ~ DaysfromTreatment)

#qqnorm(resid(model_crp_cr_INT))

```

```{r  crpmodel_cr_nonINT , echo=FALSE, include=FALSE}

# model without interaction term
model_crp_cr_nonINT <- lmer(logcrp_S_TR ~ logcrp_S_BL + DaysfromTreatment + phase + creatinine_S +(1 | ID), lmmdata_dates_MOD, na.action = na.exclude)

# summary(model_crp_cr_nonINT)
coefs_crp_cr_nonINT <- merge(
  as.data.frame(
    round(
      # getting from the model coefficients table
      # the fixed effects names --- row.names
      # the estimate --- col. 1
      # the std. error --- col. 2
      # the p-value --- col. 5
      coef(summary(model_crp_cr_nonINT))[, c(1, 2, 5)],
      digits = 3
    )
  ),
  as.data.frame(
    round(
      # getting the confidence intervals omitting the first 3 rows that
      # refer to sigma (it is NA in the Wald method)
      confint.merMod(model_crp_cr_nonINT, method = "Wald"),
      digits = 3
    )[-c(1:3), ]
  ),
  by = "row.names"
)

# change the names in the columns of the CIs
setnames(coefs_crp_cr_nonINT, old = c("Pr(>|t|)", "2.5 %", "97.5 %"), new = c("p", "CI2.5", "CI97.5"))

coefs_crp_cr_nonINT$Model <- "crp_nonadjusted_nonINT"

# revalue the fixed effects names
coefs_crp_cr_nonINT$FixedEffects <- revalue(
  coefs_crp_cr_nonINT$Row.names,
  c(
    "(Intercept)" = "Intercept",
    "DaysfromTreatment" = "Time (days)",
    "phaseOrganic" = "Organic diet",
    "logcrp_S_BL" = "Baseline crp levels (no cr adjusted)",
    "creatinine_S"="creatinine"  
  )
)

```
```{r tabmpdel Model cr, echo=FALSE, include=TRUE}
# crp
tab_model(model_crp_cr_nonINT,
  show.icc = TRUE,
  digits = 3, show.obs = TRUE, show.re.var = TRUE,
  title = paste(model_crp_cr_nonINT@call$formula[3]))

```

```{r model_crp_cr_nonINT, echo=FALSE}

plot(model_crp_cr_nonINT, resid(.) ~ fitted(.) | phase)

plot(model_crp_cr_nonINT, resid(.) ~ DaysfromTreatment)

qqnorm(resid(model_crp_cr_nonINT))

```

### Merge the models of objective 1

```{r tabmodel merge the models of objective 1, echo=FALSE, include=TRUE}
# crp
tab_model(model_crp,model_crp_cr_nonINT,
  show.icc = TRUE,
  digits = 3, show.obs = TRUE, show.re.var = TRUE,
  title = "Merge the models of objective 1")

```

### Models for objective 2

As the main analysis a second model was made to assess the association between biomarker of exposure to pesticides (3-PBA and 6-CN, separately) and biomarker of inflammation (CRP). 

 + Random part: participants
 + Fixed effect: each biomarker of exposure to pesticides, time (days of treatment, where time=0 was the start of the treatment), age and sex


### Log creatinine adjusted CRP ~ Log creatinine adjusted baseline CRP + 6-CN (binary) + DaysfromTreatment + sex + baseline age + random(...)

```{r add 6-CN, echo=FALSE, include=TRUE}
# crp

Model2<-lmer(logadj_crp_S_TR ~ logadj_crp_S_BL + cnlod + DaysfromTreatment + Sex + age_baseline_S + (1 | ID), lmmdata_dates_MOD, na.action = na.exclude)

# summary(model_crp_noINT)

coefs_model2_noINT <- merge(
  as.data.frame(
    round(
      # getting from the model coefficients table
      # the fixed effects names --- row.names
      # the estimate --- col. 1
      # the std. error --- col. 2
      # the p-value --- col. 5
      coef(summary(Model2))[, c(1, 2, 5)],
      digits = 3
    )
  ),
  as.data.frame(
    round(
      # getting the confidence intervals omitting the first 3 rows that
      # refer to sigma (it is NA in the Wald method)
      confint.merMod(Model2, method = "Wald"),
      digits = 3
    )[-c(1:3), ]
  ),
  by = "row.names"
)

# change the names in the columns of the CIs
setnames(coefs_model2_noINT, old = c("Pr(>|t|)", "2.5 %", "97.5 %"), new = c("p", "CI2.5", "CI97.5"))

coefs_model2_noINT$Model <- "crpadjusted_6CN"

# revalue the fixed effects names
coefs_model2_noINT$FixedEffects <- revalue(
  coefs_model2_noINT$Row.names,
  c(
    "cnlodAbove LOD"="6-CN (above LOD)",
    "logadj_crp_S_BL"="Baseline CRP levels",
    "DaysfromTreatment" = "Time (days)",
     "SexFemale"=="Sex (female)",
    "age_baseline_S"="Age"
  )
)

```

```{r echo=FALSE, include=TRUE}
tab_model(Model2,
  show.icc = TRUE,
  digits = 3, show.obs = TRUE, show.re.var = TRUE,
  title = paste(Model2@call$formula[3])
)

```

```{r diagnostics_model2, echo=FALSE}

plot(Model2, resid(.) ~ fitted(.) | phase)

plot(Model2, resid(.) ~ DaysfromTreatment)

qqnorm(resid(Model2))

```

### Log CRP ~ log baseline CRP + 6-CN + DaysfromTreatment + sex + baseline age + creatinine_s + random(...)

```{r add 6-CN non adjusted, echo=FALSE, include=TRUE}
# crp

Model2b<-lmer(logcrp_S_TR ~ logcrp_S_BL + cnlod + DaysfromTreatment + Sex + age_baseline_S + creatinine_S + (1 | ID), lmmdata_dates_MOD, na.action = na.exclude)

# summary(model_crp_noINT)

coefs_Model2b_noINT <- merge(
  as.data.frame(
    round(
      # getting from the model coefficients table
      # the fixed effects names --- row.names
      # the estimate --- col. 1
      # the std. error --- col. 2
      # the p-value --- col. 5
      coef(summary(Model2b))[, c(1, 2, 5)],
      digits = 3
    )
  ),
  as.data.frame(
    round(
      # getting the confidence intervals omitting the first 3 rows that
      # refer to sigma (it is NA in the Wald method)
      confint.merMod(Model2b, method = "Wald"),
      digits = 3
    )[-c(1:3), ]
  ),
  by = "row.names"
)

# change the names in the columns of the CIs
setnames(coefs_Model2b_noINT, old = c("Pr(>|t|)", "2.5 %", "97.5 %"), new = c("p", "CI2.5", "CI97.5"))

coefs_Model2b_noINT$Model <- "crp_nonadjusted_6CN"

# revalue the fixed effects names
coefs_Model2b_noINT$FixedEffects <- revalue(
  coefs_Model2b_noINT$Row.names,
  c(
    "cnlodAbove LOD"="6CN (above LOD)",
    "logcrp_S_BL"="Baseline crp levels (no cr adjusted)",
    "DaysfromTreatment" = "Time (days)",
     "SexFemale"=="Sex (female)",
    "age_baseline_S"="Age",
    "creatinine_S"="creatinine" 
    
  )
)

```

```{r echo=FALSE, include=TRUE}
tab_model(Model2b,
  show.icc = TRUE,
  digits = 3, show.obs = TRUE, show.re.var = TRUE,
  title = paste(Model2b@call$formula[3])
)

```

```{r diagnostics_Model2b, echo=FALSE}

plot(Model2b, resid(.) ~ fitted(.) | phase)

plot(Model2b, resid(.) ~ DaysfromTreatment)

qqnorm(resid(Model2b))

```

### Merge the models of objective 2-6-CN

```{r merge model of objective 2 6-cn, echo=FALSE, include=TRUE}

tab_model(Model2,Model2b,
  show.icc = TRUE,
  digits = 3, show.obs = TRUE, show.re.var = TRUE,
  title = "Merge the models of objective 2-6-CN")

```

### Log creatinine adjusted CRP ~ Log creatinine adjusted baseline CRP + creatinine adjusted 3-PBA + DaysfromTreatment + sex + baseline age + random(...)

```{r add 3-PBA, echo=FALSE, include=TRUE}
# PBA

Model3<-lmer(logadj_crp_S_TR ~ logadj_crp_S_BL + adj_pb1000_S + DaysfromTreatment + Sex + age_baseline_S + (1 | ID), lmmdata_dates_MOD, na.action = na.exclude)

# summary(model_crp_noINT)

coefs_model3_noINT <- merge(
  as.data.frame(
    round(
      # getting from the model coefficients table
      # the fixed effects names --- row.names
      # the estimate --- col. 1
      # the std. error --- col. 2
      # the p-value --- col. 5
      coef(summary(Model3))[, c(1, 2, 5)],
      digits = 3
    )
  ),
  as.data.frame(
    round(
      # getting the confidence intervals omitting the first 3 rows that
      # refer to sigma (it is NA in the Wald method)
      confint.merMod(Model3, method = "Wald"),
      digits = 3
    )[-c(1:3), ]
  ),
  by = "row.names"
)

# change the names in the columns of the CIs
setnames(coefs_model3_noINT, old = c("Pr(>|t|)", "2.5 %", "97.5 %"), new = c("p", "CI2.5", "CI97.5"))

coefs_model3_noINT$Model <- "crpadjusted_3PBA"

# revalue the fixed effects names
# revalue the fixed effects names
coefs_model3_noINT$FixedEffects <- revalue(
  coefs_model3_noINT$Row.names,
  c(
    "adj_pb1000_S"="3PBA",
    "logadj_crp_S_BL"="Baseline crp levels",
    "DaysfromTreatment" = "Time (days)",
     "SexFemale"=="Sex (female)",
    "age_baseline_S"="Age"
  )
)

```
```{r echo=FALSE, include=TRUE}
tab_model(Model3,
  show.icc = TRUE,
  digits = 3, show.obs = TRUE, show.re.var = TRUE,
  title = paste(Model3@call$formula[3])
)

```

```{r diagnostics_model3, echo=FALSE}

plot(Model3, resid(.) ~ fitted(.) | phase)

plot(Model3, resid(.) ~ DaysfromTreatment)

qqnorm(resid(Model3))

```

### Log CRP ~ Log baseline CRP + 3-PBA + DaysfromTreatment + sex + baseline age + creatinine_S + random(...)

```{r add 3-PBA non adjusted, echo=FALSE, include=TRUE}
# PBA

Model3b<-lmer(logcrp_S_TR ~ logcrp_S_BL + pb1000_S + DaysfromTreatment + Sex + age_baseline_S + creatinine_S + (1 | ID), lmmdata_dates_MOD, na.action = na.exclude)

# summary(model_crp_noINT)

coefs_Model3b_noINT <- merge(
  as.data.frame(
    round(
      # getting from the model coefficients table
      # the fixed effects names --- row.names
      # the estimate --- col. 1
      # the std. error --- col. 2
      # the p-value --- col. 5
      coef(summary(Model3b))[, c(1, 2, 5)],
      digits = 3
    )
  ),
  as.data.frame(
    round(
      # getting the confidence intervals omitting the first 3 rows that
      # refer to sigma (it is NA in the Wald method)
      confint.merMod(Model3b, method = "Wald"),
      digits = 3
    )[-c(1:3), ]
  ),
  by = "row.names"
)

# change the names in the columns of the CIs
setnames(coefs_Model3b_noINT, old = c("Pr(>|t|)", "2.5 %", "97.5 %"), new = c("p", "CI2.5", "CI97.5"))

coefs_Model3b_noINT$Model <- "crp_nonadjusted_6PBA"

# revalue the fixed effects names
# revalue the fixed effects names
coefs_Model3b_noINT$FixedEffects <- revalue(
  coefs_Model3b_noINT$Row.names,
  c(
    "pb1000_S"="3-PBA",
    "log_crp_S_BL"="Baseline CRP levels (no cr adjusted)",
    "DaysfromTreatment" = "Time (days)",
     "SexFemale"=="Sex (female)",
    "age_baseline_S"="Age",
    "creatinine_S"="creatinine"
  )
)

```

```{r echo=FALSE, include=TRUE}

tab_model(Model3b,
  show.icc = TRUE,
  digits = 3, show.obs = TRUE, show.re.var = TRUE,
  title = paste(Model3b@call$formula[3])
)

```

```{r diagnostics_Model3b, echo=FALSE}

plot(Model3b, resid(.) ~ fitted(.) | phase)

plot(Model3b, resid(.) ~ DaysfromTreatment)

qqnorm(resid(Model3b))

```

### Merge the models of objective 2-3-PBA

```{r merge the models of objective 2-3-pba, echo=FALSE, include=TRUE}

tab_model(Model3,Model3b,
  show.icc = TRUE,
  digits = 3, show.obs = TRUE, show.re.var = TRUE,
  title = "Merge the models of objective 2 3-BPA")

```

## Benjamini-Hochberg p-value adjustment - without the intercepts

```{r pvalueadj_all2, echo=FALSE, include=FALSE}

# combine the p-values of all the models made so far
pvalmod2 <- list(
  coefs_crp_INT[, c("Model", "FixedEffects", "p")],
  coefs_crp_cr_nonINT[, c("Model", "FixedEffects", "p")],
  coefs_model2_noINT[, c("Model", "FixedEffects", "p")],
  coefs_Model2b_noINT[, c("Model", "FixedEffects", "p")],
  coefs_model3_noINT[, c("Model", "FixedEffects", "p")],
  coefs_Model3b_noINT[, c("Model", "FixedEffects", "p")]
  ) # 2

all_pvalmod2 <- bind_rows(pvalmod2)

### Order data by p-value
all_pvalmod2 <- all_pvalmod2[order(all_pvalmod2$p), ]

# Benjamini-Hochberg
all_pvalmod2$BH_p <- round(p.adjust(all_pvalmod2$p, method = "BH", n = 30), digits = 3)

```

`r kable(all_pvalmod2)`

# Session information

```{r session, echo=FALSE}
sessionInfo()
```
